---
title: "PVA of the NBI: Demography - Data Exploration" ## name of your project
author: "Sinah Drenske"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 12, fig.height = 8, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Hypothesis:** We hypothesize that at the present time the northern bald ibis population can survive without further management and release. We predict that the observed demographic rates will ensure population growth and do not differ between the breeding colonies. 
* **Study area:** Austria, Germany and Italy
* **Data:** Life History data of the NBI

In this script we created different plots for the description of the NBI population. We created Timelines, Population size per year and Barplots for all individuals and the alive females to show the distribution between the sexes, stages, colonies, raising types and the mortality types.


# Setup

```{r packages}
## libraries used in this script
library("here")
library("tidyverse")
library("viridisLite")
```


# Data

```{r data}
NBI <- readRDS(file = here::here("output", "data-proc", "NBI_data.rds"))
```

Save some standard variables
```{r}
mar_default <- c(5, 4, 4, 2) + 0.1
mar_small <- c(0,0,0,0)
zero_line <- rep(0, 12)
```


# Population size per year
Plot the whole Population size per year
```{r}
# Calculate the number of living individuals per year
popsize_NBI <- tibble(
  "2008" = ifelse(test = NBI$hatch_year <= 2008 & (NBI$death_year >= 2008 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(), 
  "2009" = ifelse(test = NBI$hatch_year <= 2009 & (NBI$death_year >= 2009 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2010" = ifelse(test = NBI$hatch_year <= 2010 & (NBI$death_year >= 2010 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2011" = ifelse(test = NBI$hatch_year <= 2011 & (NBI$death_year >= 2011 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2012" = ifelse(test = NBI$hatch_year <= 2012 & (NBI$death_year >= 2012 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2013" = ifelse(test = NBI$hatch_year <= 2013 & (NBI$death_year >= 2013 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2014" = ifelse(test = NBI$hatch_year <= 2014 & (NBI$death_year >= 2014 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2015" = ifelse(test = NBI$hatch_year <= 2015 & (NBI$death_year >= 2015 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2016" = ifelse(test = NBI$hatch_year <= 2016 & (NBI$death_year >= 2016 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2017" = ifelse(test = NBI$hatch_year <= 2017 & (NBI$death_year >= 2017 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2018" = ifelse(test = NBI$hatch_year <= 2018 & (NBI$death_year >= 2018 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum(),
  "2019" = ifelse(test = NBI$hatch_year <= 2019 & (NBI$death_year >= 2019 | NBI$death_year == 0), 
                 yes = 1, 
                 no = 0) %>% 
          sum())

# Calculate the change per year
for (i in 2:ncol(popsize_NBI)) {
  
  popsize_NBI[2,i] <- popsize_NBI[1,i] - popsize_NBI[1,i-1]

}
```


```{r, fig.height=7, fig.width=10}
# only numbers
pop_num <- as.numeric(popsize_NBI[1,])

# plot it
par(mar = (c(5, 4.5, 1, 2) + 0.1))

plot(pop_num, type = "n", xaxt = "n", yaxt = "n", xlab = "Year", ylab = "Number of individuals", cex.lab = 2, ylim = c(0,200))
axis(1, at = 1:12, labels = 2008:2019, cex.axis = 2)
axis(2, cex.axis = 2)

lines(zero_line, col = "grey85")
polygon(c(1:12, rev(1:12)), c(pop_num, rev(zero_line)), col = "grey85", border= NA)

lines(pop_num, lwd = 3)
points(pop_num, pch = 16, cex = 1.5)

box()
par(mar = mar_default)
```

Save the plot
```{r}
# png(filename = here("plots", "01_description", "Popsize_per_year.png"))
# par(mar = (c(5, 4.5, 1, 2) + 0.1))
# 
# plot(pop_num, type = "n", xaxt = "n", yaxt = "n", xlab = "Year", ylab = "Number of individuals", cex.lab = 2, ylim = c(0,200))
# axis(1, at = 1:12, labels = 2008:2019, cex.axis = 2)
# axis(2, cex.axis = 2)
# 
# lines(zero_line, col = "grey85")
# polygon(c(1:12, rev(1:12)), c(pop_num, rev(zero_line)), col = "grey85", border= NA)
# 
# lines(pop_num, lwd = 3)
# points(pop_num, pch = 16, cex = 1.5)
# 
# box()
# par(mar = mar_default)
# 
# dev.off()
```

# Customized barplots
## Overall population
Show the values for each category and build a matrix with them. The NAs between some values are there for some space between the categories 

```{r}
# Comparison between delivered individuals and all other individuals
NBI_deliv <- NBI[NBI$mortality_cause == "delivered",] # 9 delivered individuals (delivered to a colony outside the project)
NBI_proj <- NBI[NBI$mortality_cause != "delivered",] # 375 not delivered individuals

# Comparison of the reached stages and the life status (alive or dead)
table(NBI_proj$life_status, NBI_proj$reached_stage)

table(NBI_deliv$life_status, NBI_deliv$reached_stage) # all alive at the time they were delivered 
```

```{r}
# List the number of individuals for different characteristics
table(NBI$reached_stage)
table(NBI$sex)
table(NBI$breeding_area)
table(NBI$raising_type)
table(NBI$mortality_cause)
```

```{r}
NBI_mat <- matrix(data = c(
             202, 74,46,62,NA,           # Stage
             184,195, 5,NA,              # Sex
             127, 89,89,22,57,NA,        # Colony (B, K, Ue, R, Allo)
             213, 162,9, NA,             # Raising type
             4,46,23,15,18,2,124,9,143)) # Mortality  type (d h c, elec, inj, pred, shot. other, unk, del, alive)
```

Plot
```{r}
par(mar = c(5,5,2,0))
barplot(NBI_mat, beside = T, ylim = c(0,220), ylab = "Number of Individuals", border = NA, 
        cex.axis = 3, cex.lab = 3,
        col = c(
                magma(4, begin = 0.25, end = 0.6), "white", # Stage
                
                plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9), # Sex
                  magma(1, begin = 0.1, alpha = 0.6),"white", # Sex
                
                inferno(5,begin =0.65,  end = 0.95, direction = -1), "white", # Colony 
                
                viridis(3, begin = 0.65, end = 0.85),"white", # Raising type
                
                gray(0:7 / 8), viridis(1, begin = 0.5))) # Mortality type

axis(1, at = c(3,7.5,12.5,17.5,24.5), labels = c("", "", "", "", ""), cex.axis = 3)
axis(1, at = c(3,7.5,12.5,17.5,24.5), labels = c("Stage", "Sex", "Colony", "Raising type", "Mortality type"), cex.axis = 3, line = 1, lwd = 0)
box()
```

Save the Plot
```{r}
# png(filename = here("plots", "01_description", "Barplot_all.png"), width = 1300, height = 835)
# par(mar = c(5,5,2,0))
# 
# barplot(NBI_mat, beside = T, ylim = c(0,220), ylab = "Number of Individuals", border = NA,
#         cex.axis = 3, cex.lab = 3,
#         col = c(
#                 magma(4, begin = 0.25, end = 0.6), "white", # Stage
# 
#                 plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9), # Sex
#                   magma(1, begin = 0.1, alpha = 0.6),"white", # Sex
# 
#                 inferno(5,begin =0.65,  end = 0.95, direction = -1), "white", # Colony
# 
#                 viridis(3, begin = 0.65, end = 0.85),"white", # Raising type
# 
#                 gray(0:7 / 8), viridis(1, begin = 0.5))) # Mortality type
# 
# axis(1, at = c(3,7.5,12.5,17.5,24.5), labels = c("", "", "", "", ""), cex.axis = 3)
# axis(1, at = c(3,7.5,12.5,17.5,24.5), labels = c("Stage", "Sex", "Colony", "Raising type", "Mortality type"), cex.axis = 3, line = 1, lwd = 0)
# box()
# 
# dev.off()
```

Legend
```{r, fig.height=6, fig.width=4}
par(mar = mar_small)
plot(1, type = "n", axes = "F", xlab = "", ylab = "") # empty plot

legend(x = "center", 
       legend = c("Stage","Juveniles","1 Year Old","2 Year Old","Adults",   "",
                  "Sex","Females","Males","Unknown",   "", 
                  "Colony","Burghausen","Kuchl","Überlingen", "Rosegg", "unassigned yet",   "", 
                  "Raising type","Foster parents","Biological Parents", "Supplementation",   "", 
                  "Mortality type","During human care","Electrocution","Injury or illness","Predation","Shot","Other", "Unknown", "Delivered", "Alive"), 
       fill = c("white",magma(4, begin = 0.25, end = 0.6),   "white", # Stage
                "white",plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9), # Sex
                        magma(1, begin = 0.1, alpha = 0.6),"white", # Sex
                "white",inferno(5,begin =0.65,  end = 0.95, direction = -1),   "white", # Colony
                "white",viridis(3, begin = 0.65, end = 0.85),         "white", # Raising type
                "white" , gray(0:7 / 8), viridis(1, begin = 0.5)), # Mortality type
       border = NA, cex = 3)

par(mar = mar_default)
```

Save the Legend
```{r}
# png(filename = here("plots", "01_description", "Barplot_all_legend.png"), width = 2174, height = 1800)
# 
# par(mar = mar_small)
# plot(1, type = "n", axes = "F", xlab = "", ylab = "") # empty plot
# 
# legend(x = "center",
#        legend = c("Stage","Juveniles","1 Year Old","2 Year Old","Adults",   "",
#                   "Sex","Females","Males","Unknown",   "",
#                   "Colony","Burghausen","Kuchl","Überlingen", "Rosegg", "unassigned yet",   "",
#                   "Raising type","Foster parents","Biological Parents", "Supplementation",   "",
#                   "Mortality type","During human care","Electrocution","Injury or illness","Predation","Shot","Other", "Unknown", "Delivered", "Alive"),
#        fill = c("white",magma(4, begin = 0.25, end = 0.6),   "white", # Stage
#                 "white",plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9), # Sex
#                         magma(1, begin = 0.1, alpha = 0.6),"white", # Sex
#                 "white",inferno(5,begin =0.65,  end = 0.95, direction = -1),   "white", # Colony
#                 "white",viridis(3, begin = 0.65, end = 0.85),         "white", # Raising type
#                 "white" , gray(0:7 / 8), viridis(1, begin = 0.5)), # Mortality type
#        border = NA, cex = 3)
# 
# par(mar = mar_default)
# 
# dev.off()
```

## Alive female individuals
```{r}
table(NBI$sex, NBI$project_status) # 74 alive females, 69 alive males = 143 alive individuals
# subset the data
NBI_f <- subset(x = NBI, subset = NBI$sex == "f" & NBI$project_status == "0")
```

```{r}
table(NBI_f$reached_stage)
table(NBI_f$breeding_area)
table(NBI_f$raising_type)
```

```{r}
NBI_mat_f <- matrix(data = c(
             37, 11,8,18,NA,           # Stage
             19, 15,28,6,6,NA,         # Colony (B, K, Ue, R, Allo)
             41, 32,1))                # Raising type
            
```

Plot
```{r}
par(mar = c(5,5,2,0))
barplot(NBI_mat_f, beside = T, ylim = c(0,50), ylab = "Number of Individuals", border = NA, 
        cex.axis = 3, cex.lab = 3,
        col = c(
                magma(4, begin = 0.25, end = 0.6), "white", # Stage
                
                inferno(5,begin =0.65,  end = 0.95, direction = -1), "white", # Colony 
                
                viridis(3, begin = 0.65, end = 0.85) # Raising type
                
                ))
axis(1, at = c(3,8.5,13.5), labels = c("", "", ""), cex.axis = 3)
axis(1, at = c(3,8.5,13.5), labels = c("Stage", "Colony", "Raising type"), cex.axis = 3, line = 1, lwd = 0)

box()
```

Save the Plot
```{r}
# png(filename = here("plots", "01_description", "Barplot_alive_f.png"), width = 1300, height = 835)
# par(mar = c(5,5,2,0))
# 
# barplot(NBI_mat_f, beside = T, ylim = c(0,50), ylab = "Number of Individuals", border = NA,
#         cex.axis = 3, cex.lab = 3,
#         col = c(
#                 magma(4, begin = 0.25, end = 0.6), "white", # Stage
# 
#                 inferno(5,begin =0.65,  end = 0.95, direction = -1), "white", # Colony
# 
#                 viridis(3, begin = 0.65, end = 0.85) # Raising type
# 
#                 ))
# axis(1, at = c(3,8.5,13.5), labels = c("", "", ""), cex.axis = 3)
# axis(1, at = c(3,8.5,13.5), labels = c("Stage", "Colony", "Raising type"), cex.axis = 3, line = 1, lwd = 0)
# 
# box()
# 
# dev.off()
```

Legend
```{r, fig.height=4, fig.width=4}
par(mar = mar_small)
plot(1, type = "n", axes = "F", xlab = "", ylab = "")
legend("center", legend = c("Stage","Juveniles","1 Year Old","2 Year Old","Adults",   "",
                            "Colony","Burghausen","Kuchl","Überlingen", "Rosegg", "unassigned yet",   "",
                            "Raising type","Foster parents","Biological Parents", "Supplementation"),
       fill = c("white",magma(4, begin = 0.25, end = 0.6),   "white",
                "white",inferno(5,begin =0.65,  end = 0.95, direction = -1),   "white",
                "white",viridis(3, begin = 0.65, end = 0.85)), border = NA, cex = 1)
par(mar = mar_default)
```

Save the legend
```{r}
# png(filename = here("plots", "01_description", "Barplot_alive_f_legend.png"), res = 130) # width = 300, height = 400
# par(mar = mar_small)
# 
# plot(1, type = "n", axes = "F", xlab = "", ylab = "")
# legend("center", legend = c("Stage","Juveniles","1 Year Old","2 Year Old","Adults",   "",
#                             "Colony","Burghausen","Kuchl","Überlingen", "Rosegg", "unassigned yet",   "",
#                             "Raising type","Foster parents","Biological Parents", "Supplementation"),
#        fill = c("white",magma(4, begin = 0.25, end = 0.6),   "white",
#                 "white",inferno(5,begin =0.65,  end = 0.95, direction = -1),   "white",
#                 "white",viridis(3, begin = 0.65, end = 0.85)), border = NA, cex = 1)
# 
# par(mar = mar_default)
# dev.off()
```

***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

