---
title: "PVA of the NBI: Demography - Survival analysis" ## name of your project
author: "Sinah Drenske, Viktoriia Radchuk"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 9, fig.height = 6, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Hypothesis:** We hypothesize that at the present time the northern bald ibis population can survive without further management and release. We predict that the observed demographic rates will ensure population growth and do not differ between the breeding colonies. 
* **Study area:** Austria, Germany and Italy
* **Data:** Life History data and Breeding data of the NBI

In this script we do the survival analysis. We calculate the survival probabilities per age class for the whole population, per raising type, per colony and per sex.

# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("here")
library("survival")
library("survminer")
library("tidyverse")
library("viridisLite")
library("timelineS")
```


# Data

```{r data}
NBI <- readRDS(file = here("output", "data-proc", "NBI_data.rds"))
NBI_breeding_all <- readRDS(file = here("output", "data-proc", "NBI_breeding.rds")) # For pre-fledgling survival

# Prepare two new columns
## The status has to be a number not a factor, otherwise the calculations are not correct
NBI$life_status_num <- as.numeric(NBI$life_status)
NBI$project_status_num <- as.numeric(NBI$project_status)
# for the paper we will consider only the **life status**
```

# Estimating survival probabilities
Procedure
1. Create survival objects with the days of life. 
   There are extra survival objects for stage, colony and sex, 
   since not all values are considered 
   (see explanations under the respective points under 3.)
2. Create different survival curves with the Kaplan-Meier estimate, 
   divided by the defined group: 
   Overall (~1), Stage, Colony, Raising type and Sex
3. Log-rank test: Test if there is a statistical difference between the survival times 
   of two or more samples
4. Plot the Kaplan-Meier estimate with the p-value from the Log-rank test
5. Cox proportional hazards model

## 1. Overall
Survival of the whole population, without further differences

Survival object
```{r overall surv obj}
# Surv() creates a survival object with the follow-up time (days of life, dof) for right-censored data and the event (life status: dead or alive). The event has to be a numeric variable, otherwise the calculation is not correct

# Life status
s_days <- Surv(time = NBI$dof, event = NBI$life_status_num)

## Project status
#s_days2 <- Surv(time = NBI$dof, event = NBI$project_status_num)
```

Survival curve
```{r overall surv curve}
# Creates survival curves from a formula. Here it is the survival object against 1, because we want to test all individuals 
# without considering subgroups.
sfit_days <- survfit(s_days ~ 1, data = NBI)
sfit_days

# sfit_days2 <- survfit(s_days2 ~ 1, data = NBI)
# sfit_days2
```


Cox proportional hazards model
```{r overall cox ph}
# Life status
# Use the created survival object on the left side of the "~" and terms you want to test on the right side
cox_overall <- coxph(Surv(NBI$dof, NBI$life_status_num) ~ 1, data = NBI)
sum_cox_overall <- summary(cox_overall) # summary of the cox proportional hazards model
sum_cox_overall

# Project status
# cox_overall2 <- coxph(Surv(NBI$dof, NBI$project_status_num) ~ 1, data = NBI)
# sum_cox_overall2 <- summary(cox_overall2) # summary of the cox proportional hazards model
# sum_cox_overall2
```


Survival plot
```{r overall surv plot, fig.height=7, fig.width=10}
ggsurvplot(sfit_days, 
           data = NBI, 
           palette = "black", 
           xlab = "Years", 
           font.x = 35, 
           font.tickslab = 34, 
           font.y = 35, 
           font.legend = 35,
           risk.table = F, 
           break.time.by = 365.25, 
           xscale = 365.25, 
           legend = "none",
           ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), 
                                        panel.grid.minor = element_blank()))
```


Save the Plot
```{r overall save plot}
# png(filename = here("plots", "02_survival", "survplot_life_overall.png"), 
#     width = 1200, height = 772)
# ggsurvplot(sfit_days, 
#            data = NBI, 
#            palette = "black", 
#            xlab = "Years", 
#            font.x = 35, 
#            font.tickslab = 34, 
#            font.y = 35, 
#            font.legend = 35,
#            risk.table = F, 
#            break.time.by = 365.25, 
#            xscale = 365.25, 
#            legend = "none",
#            ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), 
#                                         panel.grid.minor = element_blank()))
# dev.off()
```


## 2. Stages
### 2.1 All life history stages
For the whole population, but with differences between the stages

Survival object
We have to split the NBI dataset. Now every individual has as many lines as reached stages.
```{r stage surv obj}
# life status
NBI_timesplit <- survSplit(formula = Surv(NBI$dof, NBI$life_status_num) ~ ., 
                           data = NBI, 
                           cut = c(365,730,1095), 
                           episode = "STAGE")

# project status
# NBI_timesplit2 <- survSplit(formula = Surv(NBI$dof, NBI$project_status_num) ~ ., 
#                             data = NBI, 
#                             cut = c(365,730,1095), 
#                             episode = "STAGE")
```

Then create a start and end time per stage for the individuals. 
```{r stage prep obj}
# life status
## TimeStart is 0 for all rows, so that all stages start at the same time
## With this, all stages start at the same time. Otherwise, we would definitely discover significant differences, as the stages already start at different times by themselves, as one stage follows the other.
NBI_timesplit$TimeStart <- 0
## TimeStop is the difference of tstop - tstart. We take into account that the start time for all stages is 0 and not 365, 730 or 1095.  
NBI_timesplit$TimeStop <- NBI_timesplit$tstop - NBI_timesplit$tstart
NBI_timesplit$STAGE_F <- as.factor(NBI_timesplit$STAGE)


# # project status
# NBI_timesplit2$STAGE_F <- as.factor(NBI_timesplit2$STAGE)
# NBI_timesplit2$TimeStart <- 0
# ## TimeStop is the difference of tstop - tstart. With this we take into account that the start time is 0 for all lines
# NBI_timesplit2$TimeStop <- NBI_timesplit2$tstop - NBI_timesplit2$tstart
```

Survival curve
We use the curves where the survival probabilities overlap right from the start, because we do not want to take time differences into account

```{r stage surv curve life}
# life status
sfit_stage <- survfit(formula = Surv(time = NBI_timesplit$TimeStart,
                                      time2 = NBI_timesplit$TimeStop, 
                                      event = NBI_timesplit$event) ~ NBI_timesplit$STAGE_F, 
                       data = NBI_timesplit)
sum_sfit_stage <- summary(sfit_stage, times = c(0, 365, 730, 1095, 1460, 1825, 2190, 2555, 2920, 3285))
sum_sfit_stage
```


```{r surv curve proj}
# # project status
# sfit_stage2 <- survfit(formula = Surv(time = NBI_timesplit2$TimeStart,
#                                       time2 = NBI_timesplit2$TimeStop, 
#                                       event = NBI_timesplit2$event) ~ NBI_timesplit2$STAGE_F, 
#                        data = NBI_timesplit2)
# sum_sfit_stage2 <- summary(sfit_stage2, times = c(0, 365, 730, 1095, 1460, 1825, 2190, 2555, 2920, 3285))
# sum_sfit_stage2
# 
# # differences to the survival of the life status only in s1 and s2. The values for s3 and s4 are the same
```

Calculation of s4
We calculate a change rate per year in s4. This stage comprises many years, from 3 years of life on.
```{r stage s4}
# life status & project status (numbers are equal)
s4_1 <- sum_sfit_stage$surv[8]/sum_sfit_stage$surv[7] #survival change for the first year in s4, they start with a survival of 1 in each stage
s4_2 <- sum_sfit_stage$surv[9]/sum_sfit_stage$surv[8] # survival change for the second year in s4
s4_3 <- sum_sfit_stage$surv[10]/sum_sfit_stage$surv[9] 
s4_4 <- sum_sfit_stage$surv[11]/sum_sfit_stage$surv[10]
s4_5 <- sum_sfit_stage$surv[12]/sum_sfit_stage$surv[11]

s4 <- mean(c(s4_1,s4_2,s4_3,s4_4,s4_5))
s4
```
Calculate the standard deviation (SD) for the baseline values 
Formula: SE * sqrt(n) = SD                  
SE = Standard error, sqrt = Square root, n = number individuals
```{r stage sd}
# life status & project status (numbers are equal)
sd_s1 <- round(sum_sfit_stage$std.err[2]*sqrt(sum_sfit_stage$n.risk[2]), digits = 2)
sd_s2 <- round(sum_sfit_stage$std.err[4]*sqrt(sum_sfit_stage$n.risk[4]), digits = 2)
sd_s3 <- round(sum_sfit_stage$std.err[6]*sqrt(sum_sfit_stage$n.risk[6]), digits = 2)

sd_s4 <- round(sd(c(s4_1,s4_2,s4_3,s4_4,s4_5)), digits = 2)

sd_s1
sd_s2
sd_s3
sd_s4
```


Cox proportional hazards model
```{r stage cox ph}
# Life status
## Use the values where the start times and end times overlap. We do not want to take time differences into account only the survival
cox_stage <- coxph(Surv(NBI_timesplit$TimeStart, NBI_timesplit$TimeStop, NBI_timesplit$event) ~ NBI_timesplit$STAGE_F, data = NBI_timesplit)
sum_cox_stage <- summary(cox_stage)
sum_cox_stage

# # Project status
# ## Use the values where the start times and end times overlap. We do not want to take time differences into account only the survival
# cox_stage2 <- coxph(Surv(NBI_timesplit2$TimeStart2, NBI_timesplit2$TimeStop2, NBI_timesplit2$event) ~ NBI_timesplit2$STAGE_F, data = NBI_timesplit2)
# sum_cox_stage2 <- summary(cox_stage2)
# sum_cox_stage2
# 
# # No significant differences for all individuals
```

Survival plot
```{r stage surv plot, fig.height=7, fig.width=10}
# Life status
ggsurvplot(sfit_stage, data = NBI_timesplit, conf.int = T, risk.table = F, 
           xlab = "Years", xlim = c(0,5*365.25), font.x = 35, font.tickslab = 34, font.y = 35,font.legend = 35,
           ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
           break.time.by = 365.25, xscale = 365.25, pval = 0.4, pval.coord = c(x = 4.34*365.25, y = 0.8), pval.size = 8, 
           legend = c(0.9,0.89), legend.title = "Stages", legend.labs = c("Juveniles", "1 Year Old", "2 Year Old", "Adults"), 
           palette = magma(4, begin = 0.25, end = 0.6))

# # Project status
# ggsurvplot(sfit_stage2, data = NBI_timesplit, conf.int = T, risk.table = F, 
#            xlab = "Years", xlim = c(0,5*365.25), font.x = 25, font.tickslab = 24, font.y = 25,font.legend = 25,
#            ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.1, pval.coord = c(x = 4.34*365.25, y = 0.8), pval.size = 8, 
#            legend = c(0.9,0.89), legend.title = "Stages", legend.labs = c("Juveniles", "1 Year Old", "2 Year Old", "Adults"), 
#            palette = magma(4, begin = 0.25, end = 0.6))
```

Save the Plot
```{r save plot}
# # Life status
# png(filename = here("plots","02_survival", "survplot_life_stages.png"), width = 1200, height = 772)
# 
# ggsurvplot(sfit_stage, data = NBI_timesplit, conf.int = T, risk.table = F,
#            xlab = "Years", xlim = c(0,5*365.25), font.x = 35, font.tickslab = 34, font.y = 35,font.legend = 35,
#            ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.4, pval.coord = c(x = 4.2*365.25, y = 0.7), pval.size = 12,
#            legend = c(0.9,0.85), legend.title = "Stages", legend.labs = c("Juveniles", "1 Year Old", "2 Year Old", "Adults"),
#            palette = magma(4, begin = 0.25, end = 0.6))
# 
# dev.off()
#
# # Project status
# png(filename = here("plots","02_survival", "survplot_project_stages.png"), width = 1200, height = 772)
# 
# ggsurvplot(sfit_stage2, data = NBI_timesplit, conf.int = T, risk.table = F, 
#            xlab = "Years", xlim = c(0,5*365.25), font.x = 25, font.tickslab = 24, font.y = 25,font.legend = 25,
#            ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.1, pval.coord = c(x = 4.34*365.25, y = 0.8), pval.size = 8, 
#            legend = c(0.9,0.89), legend.title = "Stages", legend.labs = c("Juveniles", "1 Year Old", "2 Year Old", "Adults"), 
#            palette = magma(4, begin = 0.25, end = 0.6))
# 
# dev.off()
```

### 2.2 Pre maturity - cumulative survival
Here we calculate the survival until the NBI reach sexual maturity

Survival object
We have to split the NBI dataset. Now every individual has one line for the time before they reach sexual maturity and one line for the time after they reached it.
```{r}
NBI_timesplit_sex <- survSplit(formula = Surv(NBI$dof, NBI$life_status_num) ~ ., 
                               data = NBI, 
                               cut = c(1095), 
                               episode = "MATURITY")
```

```{r}
# life status
## TimeStart is 0 for all rows, so that all stages start at the same time. (Explanation see line 177 ff)
NBI_timesplit_sex$TimeStart <- 0
NBI_timesplit_sex$TimeStop <- NBI_timesplit_sex$tstop - NBI_timesplit_sex$tstart
NBI_timesplit_sex$MATURITY_F <- as.factor(NBI_timesplit_sex$MATURITY)
```

Survival curve
We use the curves where the survival probabilities overlap right from the start, because we do not want to take time differences into account

```{r stage prem surv curve life}
# life status
sfit_matu <- survfit(formula = Surv(time = NBI_timesplit_sex$TimeStart,
                                      time2 = NBI_timesplit_sex$TimeStop, 
                                      event = NBI_timesplit_sex$event) ~ NBI_timesplit_sex$MATURITY_F, 
                       data = NBI_timesplit_sex)

sum_sfit_matu <- summary(sfit_matu, times = c(0, 1095, 1460, 1825, 2190, 2555, 2920, 3285))
sum_sfit_matu
```

Second possibility: Multiply the survival values of the first three stages
```{r}
sum_sfit_matu$surv[2]
sum_sfit_stage$surv[2] * sum_sfit_stage$surv[4] * sum_sfit_stage$surv[6]
# Both have the same value 0.3285587 (0.33)

# Calculate the SD
round(sum_sfit_matu$std.err[2]*sqrt(sum_sfit_matu$n.risk[2]), digits = 2)
```


## 3. Colonies
We will only use two colonies. The colony in Rosegg consists only of 22 NBI and was funded 2 years ago (2018). And we do not want to calculate here the survival for individuals that do not belong to a colony yet. 
```{r col sub}
NBI_2col <- subset(NBI, NBI$breeding_area == "Burghausen" | NBI$breeding_area == "Kuchl")
NBI_2col <- droplevels(NBI_2col)
```

Survival object
```{r col surv obj}
# Life status
s_days_2col <- Surv(time = NBI_2col$dof, event = NBI_2col$life_status_num)
```

Survival curve
```{r col surv curve}
# Life status
sfit_2col <- survfit(s_days_2col ~ NBI_2col$breeding_area, data = NBI_2col)
sfit_2col
```


Cox proportional hazards model
```{r col cox ph}
# Life status
cox_2col <- coxph(Surv(time = NBI_2col$dof, event = NBI_2col$life_status_num) ~ NBI_2col$breeding_area, data = NBI_2col)
sum_cox_col <- summary(cox_2col)
sum_cox_col
```



Survival plot
```{r col surv plot}
# Life status
ggsurvplot(sfit_2col, data = NBI_3col, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
           font.x = 23, font.tickslab = 22, font.y = 23,font.legend = 22,
           ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
           break.time.by = 365.25, xscale = 365.25, pval = 0.2, pval.coord = c(x = 6.95*365.25, y = 0.85), pval.size = 7,
           legend = c(0.9,0.91), legend.title = "Colony", legend.labs = c("Burghausen", "Kuchl"), 
           palette = inferno(5,begin =0.65,  end = 0.95, direction = -1))
```

Save the Plot
```{r col save plot}
# # Life status
# png(filename = here("plots","02_survival", "survplot_life_colonies.png"), width = 1200, height = 772)
# ggsurvplot(sfit_3col, data = NBI_3col, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
#            font.x = 35, font.tickslab = 34, font.y = 35,font.legend = 35,
#            ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.2, pval.coord = c(x = 6.55*365.25, y = 0.80), pval.size = 12,
#            legend = c(0.89,0.90), legend.title = "Colony", legend.labs = c("Burghausen", "Kuchl"),
#            palette = inferno(5,begin =0.65,  end = 0.95, direction = -1))
# dev.off()
```


## 4. Colonies and stages
This is done for a better differentiation between the colonies and the stages.

Survival object
```{r sc surv obj}
# Life status
#NBI_timesplit_colony <- subset(NBI_timesplit, NBI_timesplit$breeding_area == "Burghausen" | NBI_timesplit$breeding_area == "Kuchl" | NBI_timesplit$breeding_area == "Ueberlingen")
NBI_timesplit_colony <- subset(NBI_timesplit, NBI_timesplit$breeding_area == "Burghausen" | NBI_timesplit$breeding_area == "Kuchl")
NBI_timesplit_colony <- droplevels(NBI_timesplit_colony)

# # Project status
# #NBI_timesplit_colony2 <- subset(NBI_timesplit2, NBI_timesplit2$breeding_area == "Burghausen" | NBI_timesplit2$breeding_area == "Kuchl" | NBI_timesplit2$breeding_area == "Ueberlingen")
# NBI_timesplit_colony2 <- subset(NBI_timesplit2, NBI_timesplit2$breeding_area == "Burghausen" | NBI_timesplit2$breeding_area == "Kuchl")
# NBI_timesplit_colony2 <- droplevels(NBI_timesplit_colony2)
```


Survival curve
```{r sc surv curve life}
# life status
sfit_sc <- survfit(formula = Surv(time = NBI_timesplit_colony$TimeStart2,
                                      time2 = NBI_timesplit_colony$TimeStop2, 
                                      event = NBI_timesplit_colony$event) ~ NBI_timesplit_colony$STAGE_F + NBI_timesplit_colony$breeding_area, 
                       data = NBI_timesplit_colony)
sum_sfit_sc <- summary(sfit_sc, times = c(0, 365, 730, 1095, 1460, 1825, 2190, 2555, 2920, 3285))
sum_sfit_sc
```


```{r sc surv curve proj}
# # project status
# sfit_sc2 <- survfit(formula = Surv(time = NBI_timesplit_colony2$TimeStart2,
#                                       time2 = NBI_timesplit_colony2$TimeStop2, 
#                                       event = NBI_timesplit_colony2$event) ~ NBI_timesplit_colony2$STAGE_F + NBI_timesplit_colony2$breeding_area, 
#                        data = NBI_timesplit_colony2)
# sum_sfit_sc2 <- summary(sfit_sc2, times = c(0, 365, 730, 1095, 1460, 1825, 2190, 2555, 2920, 3285))
# sum_sfit_sc2
# 
# # differences to the survival of the life status only in s1 and s2. The values for s3 and s4 are the same
```

Values of s1-s3
```{r}
# life status & project status (numbers are equal)
## Burghausen
print("Burghausen")
sum_sfit_sc$surv[2]
sum_sfit_sc$surv[6]
sum_sfit_sc$surv[10]

## Kuchl
print("Kuchl")
sum_sfit_sc$surv[4]
sum_sfit_sc$surv[8]
sum_sfit_sc$surv[12]
```

Calculation of s4 
```{r sc s4}
# life status & project status (numbers are equal)
# only for Burghausen and Kuchl - none of the individuals from Überlingen is in Stage 4 yet
## Burghausen
# s4_1_B <- sum_sfit_sc$surv[19]/sum_sfit_sc$surv[18] 
# s4_2_B <- sum_sfit_sc$surv[20]/sum_sfit_sc$surv[19] 
# s4_3_B <- sum_sfit_sc$surv[21]/sum_sfit_sc$surv[20] 
# s4_4_B <- sum_sfit_sc$surv[22]/sum_sfit_sc$surv[21]
# s4_5_B <- sum_sfit_sc$surv[23]/sum_sfit_sc$surv[22]
# 
# s4_B <- mean(c(s4_1_B,s4_2_B,s4_3_B,s4_4_B,s4_5_B))
# s4_B
# 13-18
s4_1_B <- sum_sfit_sc$surv[14]/sum_sfit_sc$surv[13] 
s4_2_B <- sum_sfit_sc$surv[15]/sum_sfit_sc$surv[14] 
s4_3_B <- sum_sfit_sc$surv[16]/sum_sfit_sc$surv[15] 
s4_4_B <- sum_sfit_sc$surv[17]/sum_sfit_sc$surv[16]
s4_5_B <- sum_sfit_sc$surv[18]/sum_sfit_sc$surv[17]

s4_B <- mean(c(s4_1_B,s4_2_B,s4_3_B,s4_4_B,s4_5_B))
s4_B

## Kuchl
# s4_1_K <- sum_sfit_sc$surv[25]/sum_sfit_sc$surv[24] 
# s4_2_K <- sum_sfit_sc$surv[26]/sum_sfit_sc$surv[25] 
# s4_3_K <- sum_sfit_sc$surv[27]/sum_sfit_sc$surv[26] 
# s4_4_K <- sum_sfit_sc$surv[28]/sum_sfit_sc$surv[27]
# s4_5_K <- sum_sfit_sc$surv[29]/sum_sfit_sc$surv[28]
# 
# s4_K <- mean(c(s4_1_K,s4_2_K,s4_3_K,s4_4_K,s4_5_K))
# s4_K
# 19-24
s4_1_K <- sum_sfit_sc$surv[20]/sum_sfit_sc$surv[19] 
s4_2_K <- sum_sfit_sc$surv[21]/sum_sfit_sc$surv[20] 
s4_3_K <- sum_sfit_sc$surv[22]/sum_sfit_sc$surv[21] 
s4_4_K <- sum_sfit_sc$surv[23]/sum_sfit_sc$surv[22]
s4_5_K <- sum_sfit_sc$surv[24]/sum_sfit_sc$surv[23]

s4_K <- mean(c(s4_1_K,s4_2_K,s4_3_K,s4_4_K,s4_5_K))
s4_K

```

Calculate the standard deviation (SD) for the baseline values 
Formula: SE * sqrt(n) = SD                  
SE = Standard error, sqrt = Square root, n = number individuals
```{r sc sd}
# life status & project status (numbers are equal)
## Burghausen
# sd_s1_B <- round(sum_sfit_sc$std.err[2]*sqrt(sum_sfit_sc$n.risk[2]), digits = 2)
# sd_s2_B <- round(sum_sfit_sc$std.err[8]*sqrt(sum_sfit_sc$n.risk[8]), digits = 2)
# sd_s3_B <- round(sum_sfit_sc$std.err[14]*sqrt(sum_sfit_sc$n.risk[14]), digits = 2)
# 
# sd_s4_B <- round(sd(c(s4_1_B,s4_2_B,s4_3_B,s4_4_B,s4_5_B)), digits = 2)
# 
# sd_s1_B
# sd_s2_B
# sd_s3_B
# sd_s4_B

sd_s1_B <- round(sum_sfit_sc$std.err[2]*sqrt(sum_sfit_sc$n.risk[2]), digits = 2)
sd_s2_B <- round(sum_sfit_sc$std.err[6]*sqrt(sum_sfit_sc$n.risk[6]), digits = 2)
sd_s3_B <- round(sum_sfit_sc$std.err[10]*sqrt(sum_sfit_sc$n.risk[10]), digits = 2)

sd_s4_B <- round(sd(c(s4_1_B,s4_2_B,s4_3_B,s4_4_B,s4_5_B)), digits = 2)

sd_s1_B
sd_s2_B
sd_s3_B
sd_s4_B


## Kuchl
# sd_s1_K <- round(sum_sfit_sc$std.err[4]*sqrt(sum_sfit_sc$n.risk[4]), digits = 2)
# sd_s2_K <- round(sum_sfit_sc$std.err[10]*sqrt(sum_sfit_sc$n.risk[10]), digits = 2)
# sd_s3_K <- round(sum_sfit_sc$std.err[16]*sqrt(sum_sfit_sc$n.risk[16]), digits = 2)
# 
# sd_s4_K <- round(sd(c(s4_1_K,s4_2_K,s4_3_K,s4_4_K,s4_5_K)), digits = 2)
# 
# sd_s1_K
# sd_s2_K
# sd_s3_K
# sd_s4_K

sd_s1_K <- round(sum_sfit_sc$std.err[4]*sqrt(sum_sfit_sc$n.risk[4]), digits = 2)
sd_s2_K <- round(sum_sfit_sc$std.err[8]*sqrt(sum_sfit_sc$n.risk[8]), digits = 2)
sd_s3_K <- round(sum_sfit_sc$std.err[12]*sqrt(sum_sfit_sc$n.risk[12]), digits = 2)

sd_s4_K <- round(sd(c(s4_1_K,s4_2_K,s4_3_K,s4_4_K,s4_5_K)), digits = 2)

sd_s1_K
sd_s2_K
sd_s3_K
sd_s4_K
```


Cox proportional hazards model
```{r sc cox ph}
# life status
cox_sc <- coxph(Surv(NBI_timesplit_colony$TimeStart2, NBI_timesplit_colony$TimeStop2, NBI_timesplit_colony$event) ~ 
                NBI_timesplit_colony$STAGE_F + NBI_timesplit_colony$breeding_area, data = NBI_timesplit_colony)
sum_cox_sc <- summary(cox_sc)
sum_cox_sc


# # Project status
# cox_sc2 <- coxph(Surv(NBI_timesplit_colony2$TimeStart2, NBI_timesplit_colony2$TimeStop2, NBI_timesplit_colony2$event) ~ 
#                 NBI_timesplit_colony2$STAGE_F + NBI_timesplit_colony2$breeding_area, data = NBI_timesplit_colony2)
# sum_cox_sc2 <- summary(cox_sc2)
# sum_cox_sc2
```


```{r sc surv plot}
# Life status
ggsurvplot(sfit_sc, data = NBI_timesplit_colony, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,5*365.25), 
           font.x = 18, font.tickslab = 18, font.y = 18,font.legend = 18,
           ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
           break.time.by = 365.25, xscale = 365.25, pval = 0.3, pval.coord = c(x = 4.25*365.25, y = 0.75), pval.size = 6, 
           legend = c(0.6,0.7), legend.title = "Colony & Stage", 
           legend.labs = c("B Stage 1","K Stage 1","B Stage 2","K Stage 2","B Stage 3", "K Stage 3","B Stage 4","K Stage 4"), 
           palette = c("#F1ED72FF", "#FCB317FF", "#F4E055FF", "#FCA50AFF", "#F7D13DFF", "#FB9606FF", "#FAC329FF", "#F8880DFF"))

# # Project status
# ggsurvplot(sfit_sc2, data = NBI_timesplit_colony2, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,5*365.25), 
#            font.x = 18, font.tickslab = 18, font.y = 18,font.legend = 18,
#            ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.08, pval.coord = c(x = 4.25*365.25, y = 0.75), pval.size = 6, 
#            legend = c(0.6,0.7), legend.title = "Colony & Stage", 
#            legend.labs = c("B Stage 1","K Stage 1","B Stage 2","K Stage 2","B Stage 3", "K Stage 3","B Stage 4","K Stage 4"), 
#            palette = c("#F1ED72FF", "#FCB317FF", "#F4E055FF", "#FCA50AFF", "#F7D13DFF", "#FB9606FF", "#FAC329FF", "#F8880DFF"))
# 
# # inferno(11,begin =0.65,  end = 0.95, direction = -1)
```


## 5. Raising types
We consider here only the raising types BP and FP, because there are only 9 individuals with the raising type "supplementation".
```{r rt sub}
NBI_RT <- subset(NBI, NBI$raising_type == "foster parent" | NBI$raising_type == "parent")
NBI_RT <- droplevels(NBI_RT)
```

Survival object
```{r rt surv obj}
# Life status
s_days_rt <- Surv(time = NBI_RT$dof, event = NBI_RT$life_status_num)

# # Project status
# s_days_rt2 <- Surv(time = NBI_RT$dof, event = NBI_RT$project_status_num)
```

Survival curve
```{r rt surv curve}
# Life status
sfit_rt <- survfit(s_days_rt ~ NBI_RT$raising_type, data = NBI_RT)
sfit_rt

# # Project status
# sfit_rt2 <- survfit(s_days_rt2 ~ NBI_RT$raising_type, data = NBI_RT)
# sfit_rt2
```


Cox proportional hazards model
```{r rt cox ph}
# Life status
cox_rt <- coxph(Surv(time = NBI_RT$dof, event = NBI_RT$life_status_num) ~ NBI_RT$raising_type, data = NBI_RT)
sum_cox_rt <- summary(cox_rt)
sum_cox_rt

# # Project status
# cox_rt2 <- coxph(Surv(time = NBI_RT$dof, event = NBI_RT$project_status_num) ~ NBI_RT$raising_type, data = NBI_RT)
# sum_cox_rt2 <- summary(cox_rt2)
# sum_cox_rt2
# 
# # Significant differences between the raising types for both options. 
```


Survival plot
```{r rt surv plot}
# Life status
ggsurvplot(sfit_rt, data = NBI_RT, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
           font.x = 23, font.tickslab = 22, font.y = 23,font.legend = 22,
           ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
           break.time.by = 365.25, xscale = 365.25, pval = 0.004, pval.coord = c(x = 6.75*365.25, y = 0.85), pval.size = 7,
           legend = c(0.9,0.91), legend.title = "Raising type", 
           legend.labs = c("Foster parents", "Biological parents"), 
           palette = viridis(3, begin = 0.65, end = 0.85))

# # Project status
# ggsurvplot(sfit_rt2, data = NBI_RT, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
#            font.x = 23, font.tickslab = 22, font.y = 23,font.legend = 22,
#            ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = T, pval.coord = c(x = 6.75*365.25, y = 0.85), pval.size = 7,
#            legend = c(0.9,0.91), legend.title = "Raising type", 
#            legend.labs = c("Foster parents", "Biological parents"), 
#            palette = viridis(3, begin = 0.65, end = 0.85))
```

Save the Plot
```{r rt save plot}
# Life status
# png(filename = here("plots","02_survival", "survplot_life_raising.png"), width = 1200, height = 772)
# 
# ggsurvplot(sfit_rt, data = NBI_RT, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
#            font.x = 35, font.tickslab = 34, font.y = 35,font.legend = 35,
#            ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.004, pval.coord = c(x = 5.85*365.25, y = 0.8), pval.size = 12,
#            legend = c(0.85,0.90), legend.title = "Raising type",
#            legend.labs = c("Foster parents", "Biological parents"),
#            palette = viridis(3, begin = 0.65, end = 0.85))
# 
# dev.off()
#
# # Project status
# png(filename = here("plots","02_survival", "survplot_project_raising.png"), width = 1200, height = 772)
# 
# ggsurvplot(sfit_rt2, data = NBI_RT, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
#            font.x = 23, font.tickslab = 22, font.y = 23,font.legend = 22,
#            ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.0005, pval.coord = c(x = 6.75*365.25, y = 0.85), pval.size = 7,
#            legend = c(0.9,0.91), legend.title = "Raising type", 
#            legend.labs = c("Foster parents", "Biological parents"), 
#            palette = viridis(3, begin = 0.65, end = 0.85))
# 
# dev.off()
```


## 6. Raising types and stages

Survival object
```{r srt surv obj}
# Life status
NBI_timesplit_srt <- subset(NBI_timesplit, NBI_timesplit$raising_type == "foster parent" | NBI_timesplit$raising_type == "parent")
NBI_timesplit_srt <- droplevels(NBI_timesplit_srt)

# # Project status
# NBI_timesplit_srt2 <- subset(NBI_timesplit2, NBI_timesplit2$raising_type == "foster parent" | NBI_timesplit2$raising_type == "parent")
# NBI_timesplit_srt2 <- droplevels(NBI_timesplit_srt2)
```


Survival curve
```{r srt surv curve life}
# life status
sfit_srt <- survfit(formula = Surv(time = NBI_timesplit_srt$TimeStart2,
                                      time2 = NBI_timesplit_srt$TimeStop2, 
                                      event = NBI_timesplit_srt$event) ~ NBI_timesplit_srt$STAGE_F + NBI_timesplit_srt$raising_type, 
                       data = NBI_timesplit_srt)
sum_sfit_srt <- summary(sfit_srt, times = c(0, 365, 730, 1095, 1460, 1825, 2190, 2555, 2920, 3285))
sum_sfit_srt
```


```{r srt surv curve proj}
# # project status
# sfit_srt2 <- survfit(formula = Surv(time = NBI_timesplit_srt2$TimeStart2,
#                                       time2 = NBI_timesplit_srt2$TimeStop2, 
#                                       event = NBI_timesplit_srt2$event) ~ NBI_timesplit_srt2$STAGE_F + NBI_timesplit_srt2$raising_type, 
#                        data = NBI_timesplit_srt2)
# sum_sfit_srt2 <- summary(sfit_srt2, times = c(0, 365, 730, 1095, 1460, 1825, 2190, 2555, 2920, 3285))
# sum_sfit_srt2
# 
# # differences to the survival of the life status only in s1 (BP) and s2 (FP). The values for s3 and s4 are the same
```

Values of s1-s3
```{r}
# life status & project status (numbers are equal)
## FP individuals
print("FP")
sum_sfit_srt$surv[2]
sum_sfit_srt$surv[6]
sum_sfit_srt$surv[10]

## BP individuals
print("BP")
sum_sfit_srt$surv[4]
sum_sfit_srt$surv[8]
sum_sfit_srt$surv[12]
```


Calculation of s4 
```{r srt s4}
# life status & project status (numbers are equal)
## FP individuals
s4_1_FP <- sum_sfit_srt$surv[14]/sum_sfit_srt$surv[13] 
s4_2_FP <- sum_sfit_srt$surv[15]/sum_sfit_srt$surv[14] 
s4_3_FP <- sum_sfit_srt$surv[16]/sum_sfit_srt$surv[15] 
s4_4_FP <- sum_sfit_srt$surv[17]/sum_sfit_srt$surv[16]
s4_5_FP <- sum_sfit_srt$surv[18]/sum_sfit_srt$surv[17]

s4_FP <- mean(c(s4_1_FP,s4_2_FP,s4_3_FP,s4_4_FP,s4_5_FP))
s4_FP

## BP individuals
s4_1_BP <- sum_sfit_srt$surv[20]/sum_sfit_srt$surv[19] 
s4_2_BP <- sum_sfit_srt$surv[21]/sum_sfit_srt$surv[20] 
s4_3_BP <- sum_sfit_srt$surv[22]/sum_sfit_srt$surv[21] 
s4_4_BP <- sum_sfit_srt$surv[23]/sum_sfit_srt$surv[22]
s4_5_BP <- sum_sfit_srt$surv[24]/sum_sfit_srt$surv[23]

s4_BP <- mean(c(s4_1_BP,s4_2_BP,s4_3_BP,s4_4_BP,s4_5_BP))
s4_BP

```

Calculate the standard deviation (SD) for the baseline values 
Formula: SE * sqrt(n) = SD                  
SE = Standard error, sqrt = Square root, n = number individuals
```{r srt sd}
# life status & project status (numbers are equal)
## FP individuals
sd_s1_FP <- round(sum_sfit_srt$std.err[2]*sqrt(sum_sfit_srt$n.risk[2]), digits = 2)
sd_s2_FP <- round(sum_sfit_srt$std.err[6]*sqrt(sum_sfit_srt$n.risk[6]), digits = 2)
sd_s3_FP <- round(sum_sfit_srt$std.err[10]*sqrt(sum_sfit_srt$n.risk[10]), digits = 2)

sd_s4_FP <- round(sd(c(s4_1_FP,s4_2_FP,s4_3_FP,s4_4_FP,s4_5_FP)), digits = 2)

print("FP")
sd_s1_FP
sd_s2_FP
sd_s3_FP
sd_s4_FP

## BP individuals
sd_s1_BP <- round(sum_sfit_srt$std.err[4]*sqrt(sum_sfit_srt$n.risk[4]), digits = 2)
sd_s2_BP <- round(sum_sfit_srt$std.err[8]*sqrt(sum_sfit_srt$n.risk[8]), digits = 2)
sd_s3_BP <- round(sum_sfit_srt$std.err[12]*sqrt(sum_sfit_srt$n.risk[12]), digits = 2)

sd_s4_BP <- round(sd(c(s4_1_BP,s4_2_BP,s4_3_BP,s4_4_BP,s4_5_BP)), digits = 2)

print("BP")
sd_s1_BP
sd_s2_BP
sd_s3_BP
sd_s4_BP
```


Cox proportional hazards model
```{r srt cox ph}
# life status
cox_srt <- coxph(Surv(NBI_timesplit_srt$TimeStart2, NBI_timesplit_srt$TimeStop2, NBI_timesplit_srt$event) ~ 
                NBI_timesplit_srt$STAGE_F + NBI_timesplit_srt$raising_type, data = NBI_timesplit_srt)
sum_cox_srt <- summary(cox_srt)
sum_cox_srt


# # Project status
# cox_srt2 <- coxph(Surv(NBI_timesplit_srt2$TimeStart2, NBI_timesplit_srt2$TimeStop2, NBI_timesplit_srt2$event) ~ 
#                 NBI_timesplit_srt2$STAGE_F + NBI_timesplit_srt2$raising_type, data = NBI_timesplit_srt2)
# sum_cox_srt2 <- summary(cox_srt2)
# sum_cox_srt2
```


Survival plot
```{r srt surv plot}
# Life status
ggsurvplot(sfit_srt, data = NBI_timesplit_srt, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,5*365.25), 
           font.x = 18, font.tickslab = 18, font.y = 18,font.legend = 18,
           ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
           break.time.by = 365.25, xscale = 365.25, pval = 0.04, pval.coord = c(x = 4.25*365.25, y = 0.75), pval.size = 6, 
           legend = c(0.6,0.7), legend.title = "Colony & Stage", 
           legend.labs = c("FP Stage 1","BP Stage 1","FP Stage 2","BP Stage 2", "FP Stage 3", "BP Stage 3", "FP Stage 4","BP Stage 4"), 
           palette = c("#2FB47CFF","#65CB5EFF","#3ABA76FF","#75D054FF","#47C06FFF","#87D549FF","#55C668FF","#9AD93CFF"))

# # Project status
# ggsurvplot(sfit_srt2, data = NBI_timesplit_srt2, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,5*365.25), 
#            font.x = 18, font.tickslab = 18, font.y = 18,font.legend = 18,
#            ggtheme = theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.005, pval.coord = c(x = 4.25*365.25, y = 0.75), pval.size = 6, 
#            legend = c(0.6,0.7), legend.title = "Colony & Stage", 
#            legend.labs = c("FP Stage 1","BP Stage 1","FP Stage 2","BP Stage 2", "FP Stage 3", "BP Stage 3", "FP Stage 4","BP Stage 4"), 
#            palette = c("#2FB47CFF","#65CB5EFF","#3ABA76FF","#75D054FF","#47C06FFF","#87D549FF","#55C668FF","#9AD93CFF"))
# 
# # viridis(8, begin = 0.65, end = 0.85)
```


## 7. Sexes
To calculate differences in survival between males and females. Individuals with unknown sex are not taken into account.

```{r sex sub}
NBI_sex <- subset(NBI, NBI$sex == "f" | NBI$sex == "m")
NBI_sex <- droplevels(NBI_sex)
```

Survival object
```{r sex surv obj}
# Life status
s_days_sex <- Surv(time = NBI_sex$dof, event = NBI_sex$life_status_num)

# # Project status
# s_days_sex2 <- Surv(time = NBI_sex$dof, event = NBI_sex$project_status_num)
```

Survival curve
```{r sex surv curve}
# Life status
sfit_sex <- survfit(s_days_sex ~ NBI_sex$sex, data = NBI_sex)
sfit_sex

# # Project status
# sfit_sex2 <- survfit(s_days_sex2 ~ NBI_sex$sex, data = NBI_sex)
# sfit_sex2
```


Cox proportional hazards model
```{r sex cox ph}
# Life status
cox_sex <- coxph(Surv(time = NBI_sex$dof, event = NBI_sex$life_status_num) ~ NBI_sex$sex, data = NBI_sex)
sum_cox_sex <- summary(cox_sex)
sum_cox_sex

# # Project status
# cox_sex2 <- coxph(Surv(time = NBI_sex$dof, event = NBI_sex$project_status_num) ~ NBI_sex$sex, data = NBI_sex)
# sum_cox_sex2 <- summary(cox_sex2)
# sum_cox_sex2
```


Survival plot
```{r sex surv plot}
# Life status
ggsurvplot(sfit_sex, data = NBI_sex, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
           font.x = 23, font.tickslab = 22, font.y = 23,font.legend = 22,
           ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
           break.time.by = 365.25, xscale = 365.25, pval = 0.4, pval.coord = c(x = 6.75*365.25, y = 0.85), pval.size = 7,
           legend = c(0.9,0.91), legend.title = "Sex", 
           legend.labs = c("females", "males"), 
           palette = c(plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9)))

# # Project status
# ggsurvplot(sfit_sex2, data = NBI_sex, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
#            font.x = 23, font.tickslab = 22, font.y = 23,font.legend = 22,
#            ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = T, pval.coord = c(x = 6.75*365.25, y = 0.85), pval.size = 7,
#            legend = c(0.9,0.91), legend.title = "Sex", 
#            legend.labs = c("females", "males"), 
#            palette = c(plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9)))
```


Save the Plot
```{r sex save plot}
# Life status
png(filename = here("plots","02_survival", "survplot_life_sex.png"), width = 1200, height = 772)

ggsurvplot(sfit_sex, data = NBI_sex, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
           font.x = 35, font.tickslab = 34, font.y = 35,font.legend = 35,
           ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
           break.time.by = 365.25, xscale = 365.25, pval = 0.4, pval.coord = c(x = 7.05*365.25, y = 0.80), pval.size = 12,
           legend = c(0.92,0.90), legend.title = "Sex",
           legend.labs = c("females", "males"),
           palette = c(plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9)))

dev.off()
#
# # Project status
# png(filename = here("plots","02_survival", "survplot_project_sex.png"), width = 1200, height = 772)
# 
# ggsurvplot(sfit_sex2, data = NBI_sex, conf.int = T, risk.table = F, xlab = "Years", xlim = c(0,8*365.25),
#            font.x = 23, font.tickslab = 22, font.y = 23,font.legend = 22,
#            ggtheme = theme_bw()+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
#            break.time.by = 365.25, xscale = 365.25, pval = 0.4, pval.coord = c(x = 7.1*365.25, y = 0.85), pval.size = 7,
#            legend = c(0.9,0.91), legend.title = "Sex", 
#            legend.labs = c("females", "males"), 
#            palette = c(plasma(1, direction = -1, begin = 0.0, end = 0.6), viridis(1, begin = 0.2, alpha = 0.9)))
# 
# dev.off()
```


# Calculate survival improvements of values for life status
Build a Vector for the Baseline survival
```{r setup baseline life}
survival_Baseline <- round(c(
                              sum_sfit_stage$surv[2], # Juveniles
                              sum_sfit_stage$surv[4], # 1-Year-old NBI
                              sum_sfit_stage$surv[6], # 2-Year-old NBI
                              s4), digits = 2)         # Adults, calculated above
survival_Baseline
```

Calculate the improvements of the survival
```{r calc impro life}
# 10%
## survival probability + (survival probability / 10)
survival_10 <- round(survival_Baseline + (survival_Baseline / 10), digits = 2)
survival_10

# 25%
## survival probability + (survival probability / 4)
survival_25 <- round(survival_Baseline + (survival_Baseline / 4), digits = 2)
survival_25
```

Set up a table for the survival probabilities
```{r baseline surv table life}
survival_life <- data.frame(s1 = NA, s2 = NA, s3=NA, s4=NA)
survival_life[1,1:4] <- c(survival_Baseline)
survival_life[2,1:4] <- c(survival_10)  
survival_life[3,1:4] <- c(survival_25)

colnames(survival_life) <- c("s1","s2","s3","s4")
row.names(survival_life) <- c("Baseline","Percent_10","Percent_25")
survival_life
```

Calculate the mortality  for NetLogo
```{r baseline mort table life}
mortality_life <- 1-survival_life
mortality_life
```


# Timelines
We create a figure with one line per reached stage per individual, according to the survival analysis. 
We consider here the life status

## Preparation
Factorize the stage
```{r TL factor stage}
NBI_timesplit$STAGE <- as.factor(NBI_timesplit$STAGE)
```

Build a column for the start date (and not the days) for each stage. 
If Stage = 1, leave the hatching date. If Stage = 2, hatching date + 365 days, If stage = 4, hatching days + 1095 days. 
```{r TL start dates}
NBI_timesplit$Year <- ifelse(test = NBI_timesplit$STAGE == 1, 
                             yes = NBI_timesplit$hatching_date, 
                             no = ifelse(test = NBI_timesplit$STAGE == 2, 
                                         yes = NBI_timesplit$hatching_date + 365, 
                                         no = ifelse(test = NBI_timesplit$STAGE == 3, 
                                                     yes = NBI_timesplit$hatching_date + 730, 
                                                     no = ifelse(test = NBI_timesplit$STAGE == 4, 
                                                                 yes = NBI_timesplit$hatching_date + 1095, 
                                                                 no = as.Date("2001-01-01")))))

# Change the datetype
NBI_timesplit$Year <- as.Date(NBI_timesplit$Year, origin = "1970-01-01")
```


Build a column for the end date. Same like above, but here is the difference between dead and alive individuals important.
```{r TL end date}
NBI_timesplit$end_stage <- ifelse(
                             test = NBI_timesplit$event == 1 & NBI_timesplit$STAGE == 1, 
                             yes = NBI_timesplit$delivered_or_death_date, 
                             no = ifelse(
                                    test = NBI_timesplit$event == 0 & NBI_timesplit$STAGE == 1 , 
                                    yes = NBI_timesplit$hatching_date + 365, 
                                    no = ifelse(
                                           test = NBI_timesplit$event == 1 & NBI_timesplit$STAGE == 2, 
                                           yes = NBI_timesplit$delivered_or_death_date, 
                                           no = ifelse(
                                                  test = NBI_timesplit$event == 0 & NBI_timesplit$STAGE == 2, 
                                                  yes =  NBI_timesplit$hatching_date + 730, 
                                                  no = ifelse(
                                                         test = NBI_timesplit$event == 1 & NBI_timesplit$STAGE == 3, 
                                                         yes =  NBI_timesplit$delivered_or_death_date, 
                                                         no = ifelse(
                                                                test = NBI_timesplit$event == 0 & NBI_timesplit$STAGE == 3, 
                                                                yes =  NBI_timesplit$hatching_date + 1095, 
                                                                no = ifelse(
                                                                       test = NBI_timesplit$event == 1 & NBI_timesplit$STAGE == 4, 
                                                                       yes =  NBI_timesplit$delivered_or_death_date, 
                                                                       no = ifelse(
                                                                              test = NBI_timesplit$event == 0 & NBI_timesplit$STAGE == 4, 
                                                                              yes =  NBI_timesplit$delivered_or_death_date, 
                                                                              no = as.Date("2001-01-01")))))))))
# Change the date type
NBI_timesplit$end_stage <- as.Date(NBI_timesplit$end_stage, origin = "1970-01-01")
# Also it is important to cut down the end date on the 19.01.2018 if it is to far in the future (because of the addition of days). And to change the date type.
NBI_timesplit$end_stage <- ifelse(NBI_timesplit$end_stage > "2020-01-01", as.Date("2020-01-01"), NBI_timesplit$end_stage)
NBI_timesplit$end_stage <- as.Date(NBI_timesplit$end_stage, origin = "1970-01-01")
```


## Write a function for the plot of timelines. 
Based on the timelineG function from the timelineS package. I changed only the colors.
```{r TL function}
timeline_NBI <- function (df, start, end, names, phase = NA, group1 = NA, group2 = NA, 
  width = 2, color = "black", theme = NULL, other = NULL) 
{
  if (!is.data.frame(df)) {
    stop("'df' must be a data frame")
  }
  if (!(class(df[[start]]) == "Date") && (class(df[[end]]) == 
    "Date")) {
    stop("'start' and 'end' must be dates")
  }
  facets <- if (is.na(group1) & is.na(group2)) {
    paste(names, "~.")
  }
  else if (is.na(group2)) {
    paste(group1, "~.")
  }
  else {
    paste(group1, "~", group2)
  }
  a <- if (is.na(phase)) {
    ggplot(df, aes_string(x = start, y = names)) + geom_segment(aes_string(x = start, 
      xend = end, y = names, yend = names), size = width, color = color)
  }
  else {
    ggplot(df, aes_string(x = start, y = names, color = phase)) + 
      geom_segment(aes_string(x = start, xend = end, y = names, yend = names), size = width) +
        scale_color_manual(values=c("#51127CFF", "#7F2582FF", "#AE347BFF","#DE4968FF"))
  }
  a <- if (is.na(group1) & is.na(group2)) {
    a
  }
  else {
    a + facet_grid(facets, scales = "free_y", space = "free_y", 
      drop = TRUE)
  }
  a <- a + theme + other
  plot(a)
}
```


## Plot
```{r}
test <- NBI_timesplit[1:5,]
```

Timelines for all individuals with the stages
```{r TL plot}
timeline_NBI(df = NBI_timesplit, start = "Year", end = "end_stage", names = "name", phase = "STAGE", group1 = "hatch_year", 
             theme = theme_bw() + theme(panel.border = element_blank(), 
                                        panel.spacing.y = unit(0.0, "lines"), 
                                        panel.grid.major.y = element_blank(), 
                                        panel.grid.major.x = element_line(), 
                                        panel.grid.minor = element_line(), 
                                        strip.background = element_blank(), 
                                        strip.text = element_blank(), 
                                        axis.text.y = element_blank(), 
                                        axis.ticks.y = element_blank(), 
                                        axis.title = element_text(size = 18), 
                                        axis.text = element_text(size = 16), 
                                        legend.text = element_text(size = 16), 
                                        legend.title = element_text(size = 18)), 
             other = labs(colour = "Stage", y = "Individuals"))
```


Save the Plot
```{r TL save plot}
# png(filename = here("plots", "02_survival", "timelines_all.png"), width = 700, height = 763)
# timeline_NBI(df = NBI_timesplit, start = "Year", end = "end_stage", names = "name", phase = "STAGE", group1 = "hatch_year", 
#              theme = theme_bw() + theme(panel.border = element_blank(), 
#                                         panel.spacing.y = unit(0.0, "lines"), 
#                                         panel.grid.major.y = element_blank(), 
#                                         panel.grid.major.x = element_line(), 
#                                         panel.grid.minor = element_line(), 
#                                         strip.background = element_blank(), 
#                                         strip.text = element_blank(), 
#                                         axis.text.y = element_blank(), 
#                                         axis.ticks.y = element_blank(), 
#                                         axis.title = element_text(size = 18), 
#                                         axis.text = element_text(size = 16), 
#                                         legend.text = element_text(size = 16), 
#                                         legend.title = element_text(size = 18)), 
#              other = labs(colour = "Stage", y = "Individuals"))
# dev.off()
```

# Pre-fledgling survival
```{r}
NBI_breeding_all$breed_year <- substring(NBI_breeding_all$b_pair_date, 1, 4)

# Please note: These values differed from those used to calculate the reproduction rates of baseline, status quo and all chicks. 
# In this table, the actual nests are counted, not the individuals that could theoretically breed. In addition, nests of temporarily added individuals are included in this table.

# Add the raising type of the mothers
nest_numbers_fm <- left_join(x = NBI_breeding_all, y =  NBI[c(2,7)], by = c("b_parent_f" = "name"))

# If the raising type is missing, add the raising type "added" for temporarily added female individuals
# add a new factor level
levels(nest_numbers_fm$raising_type) <- c(levels(nest_numbers_fm$raising_type), "added")
# replace the NAs
nest_numbers_fm$raising_type <- replace_na(data = nest_numbers_fm$raising_type, replace = "added")
# remove unneeded factor levels (here supplementation)
nest_numbers_fm$raising_type <- droplevels(nest_numbers_fm$raising_type)
levels(nest_numbers_fm$raising_type)
```


```{r}
N_nests <- nest_numbers_fm %>% count(breed_year)

NBI_breeding_year <- nest_numbers_fm %>% group_by(breed_year) %>% summarize_if(is.numeric, sum, na.rm = TRUE)
NBI_breeding_year <- NBI_breeding_year[,-c(2,3)]
NBI_breeding_year$n_nests <- N_nests$n
```

```{r}
NBI_breeding_SD <- nest_numbers_fm %>% group_by(breed_year) %>% mutate(SD_eggs = sd(b_count_eggs),SD_chicks = sd(b_count_hatching), SD_fledg = sd(b_count_fledgling)) 
NBI_breeding_SD2 <- NBI_breeding_SD %>% group_by(breed_year) %>% summarize(SD_eggs = mean(SD_eggs, na.rm = T), SD_chicks = mean(SD_chicks, na.rm = T), SD_fledg = mean(SD_fledg, na.rm = T))
#NBI_breeding_SD2 <- NBI_breeding_SD2[, -c(2:6)]
```


```{r}
# Values per year

## Mean eggs per nest
NBI_breeding_year$eggs_per_nest <- NBI_breeding_year$b_count_eggs/NBI_breeding_year$n_nests
NBI_breeding_year$eggs_per_nest <- NBI_breeding_year$b_count_eggs/NBI_breeding_year$n_nests

## mean chicks per nest
NBI_breeding_year$chicks_per_nest <- NBI_breeding_year$b_count_hatching/NBI_breeding_year$n_nests

## mean fledglings per nest
NBI_breeding_year$fledg_per_nest <- NBI_breeding_year$b_count_fledgling/NBI_breeding_year$n_nests

## survival egg to chick
NBI_breeding_year$surv_egg_chick <- NBI_breeding_year$b_count_hatching/NBI_breeding_year$b_count_eggs

## survival egg to fledgling
NBI_breeding_year$surv_egg_fledg <- NBI_breeding_year$b_count_fledgling/NBI_breeding_year$b_count_eggs

## survival chick to fledgling
NBI_breeding_year$surv_chick_fledg <- NBI_breeding_year$b_count_fledgling/NBI_breeding_year$b_count_hatching
```

```{r}
# For all years
## Sum of eggs, chicks, fledglings, nestst
print("Sum")
sum(NBI_breeding_year$b_count_eggs)
sum(NBI_breeding_year$b_count_hatching)
sum(NBI_breeding_year$b_count_fledgling)
sum(NBI_breeding_year$n_nests)

## Mean number of eggs, chicks, fledglings
print("Mean")
mean(NBI_breeding_year$eggs_per_nest)
mean(NBI_breeding_year$chicks_per_nest)
mean(NBI_breeding_year$fledg_per_nest)

## SD of mean number of eggs, chicks, fledglings
print("SD")
sd(NBI_breeding_year$eggs_per_nest)
sd(NBI_breeding_year$chicks_per_nest)
sd(NBI_breeding_year$fledg_per_nest)

## Survival egg to chick
print("survival")
sum(NBI_breeding_year$b_count_hatching)/sum(NBI_breeding_year$b_count_eggs)
## Survival egg to fledgling
sum(NBI_breeding_year$b_count_fledgling)/sum(NBI_breeding_year$b_count_eggs)
## Survival chick to fledgling
sum(NBI_breeding_year$b_count_fledgling)/sum(NBI_breeding_year$b_count_hatching)

print("survival possibility 2")
#mean(NBI_breeding_year$surv_egg_chick)
sd(NBI_breeding_year$surv_egg_chick)

#mean(NBI_breeding_year$surv_egg_fledg)
sd(NBI_breeding_year$surv_egg_fledg)

#mean(NBI_breeding_year$surv_chick_fledg)
sd(NBI_breeding_year$surv_chick_fledg)

```

## Per raising type
```{r}
N_nests_raising <- nest_numbers_fm %>% count(raising_type, breed_year)

NBI_breeding_raising <- nest_numbers_fm %>% group_by(raising_type, breed_year) %>% summarize_if(is.numeric, sum, na.rm = TRUE)
NBI_breeding_raising <- NBI_breeding_raising[,-c(3,4)]
NBI_breeding_raising$n_nests <- N_nests_raising$n

# Calculate the number of fledglings per nest per year per raising type
NBI_breeding_raising$fledg_nest <- NBI_breeding_raising$b_count_fledgling/NBI_breeding_raising$n_nests
# Calculate the mean per raising type
aggregate(x = NBI_breeding_raising, by = list(Raising = NBI_breeding_raising$raising_type), FUN = mean)
round(sd(c(NBI_breeding_raising$fledg_nest[1:8])), digits = 2) # FP +-0.69
round(sd(c(NBI_breeding_raising$fledg_nest[9:11])), digits = 2) # BP +- 2.52
round(sd(c(NBI_breeding_raising$fledg_nest[12:20])), digits = 2) # added +- 0.89
```

## Per colony
```{r}
N_nests_col <- nest_numbers_fm %>% count(b_breeding_colony, breed_year)

NBI_breeding_col <- nest_numbers_fm %>% group_by(b_breeding_colony, breed_year) %>% summarize_if(is.numeric, sum, na.rm = TRUE)
NBI_breeding_col <- NBI_breeding_col[,-c(2,3)]
NBI_breeding_col$n_nests <- N_nests_col$n

# Calculate the number of fledglings per nest per year per raising type
NBI_breeding_col$fledg_nest <- NBI_breeding_col$b_count_fledgling/NBI_breeding_col$n_nests
# Calculate the mean per raising type
aggregate(x = NBI_breeding_col, by = list(Raising = NBI_breeding_col$b_breeding_colony), FUN = mean)
round(sd(c(NBI_breeding_raising$fledg_nest[1:9])), digits = 2) # Burghausen +- 0.96
round(sd(c(NBI_breeding_raising$fledg_nest[10:15])), digits = 2) # Kuchl +- 1.63
```

## Save data
```{r}
# saveRDS(object = nest_numbers_fm, file = here::here("output", "data-proc", "nest_numbers_fm.rds"))
# saveRDS(object = NBI_breeding_year, file = here::here("output", "data-proc", "NBI_breeding_year.rds"))
```

***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

