---
title: "PVA of the NBI: Demography - GLMM of the number of fledglings per potential mother and per nest" ## name of your project
author: "Sinah Drenske, Viktoriia Radchuk, Stephanie Kramer-Schadt"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 9, fig.height = 6, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Hypothesis:** We hypothesize that at the present time the northern bald ibis population can survive without further management and release. We predict that the observed demographic rates will ensure population growth and do not differ between the breeding colonies.
* **Study area:** Austria, Germany and Italy
* **Data:** Data of the number of fledglings per potential mother and per nest


We create GLMMs to investigate whether the raising type or colony has an influence on a) the number of fledglings per potential mother (incl. breeding probability) and b) the number of fledglings per nest (only actual breeding events and incl. temporarily added females as mothers).

# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("DHARMa")
library("effects")
library("ggeffects")
library("ggplot2")
library("here")
library("lme4")
library("pwr")
library("tidyverse")
```


# Data

```{r data} 
f_fledg_mother <- readRDS(file = here("output", "data-proc", "f_fledg_mother.rds")) # from script "05a_Reproduction_analysis"          # chicks_year 
f_fledg_mother <- droplevels(f_fledg_mother)
nest_numbers_fm <- readRDS(file = here("output", "data-proc", "nest_numbers_fm.rds")) # from script "04_Survival_analysis"        # NBI_breeding_all
nest_numbers_fm <- droplevels(nest_numbers_fm)
```

# Female fledglings per potential mother per year
Based on the values used for RR Baseline

```{r}
names(f_fledg_mother)
```

## GLMM of the number of fledglings per potential mother per raising type
### Data exploration
```{r}
# Plot the number of fledglings
ggplot(data = f_fledg_mother, mapping = aes(x = nr_f_fledg)) + 
  geom_histogram()

# Plot of the number of fledglings per raising type per year
f_fledg_mother %>%
  group_by(raising_type, hatch_year) %>%
  mutate(sum.fledg = sum(nr_f_fledg)) %>%
  ggplot(mapping = aes(x = hatch_year, y = sum.fledg)) + 
  geom_line(aes(group = raising_type, color = raising_type))

# Sum of the breeding events per raising type
table(f_fledg_mother$raising_type)

# Mean per group
RT_year <- aggregate(x = f_fledg_mother, by = list(Raising = f_fledg_mother$raising_type, Year = f_fledg_mother$hatch_year), 
                     FUN = mean)
RT_mean <- aggregate(x = RT_year, by = list(RT_year$Raising), 
            FUN = mean)
```

### GLM
```{r}
null_model_f_fledg_raising <- glm(formula = nr_f_fledg ~ 1, 
                                  data = f_fledg_mother, 
                                  family = "poisson")
summary(null_model_f_fledg_raising)

f_fledg_raising_glm <- glm(formula = nr_f_fledg ~ raising_type,
                           data = f_fledg_mother,
                           family = "poisson")
summary(f_fledg_raising_glm) # non significant
```

#### Plot
```{r}
#dredge(raising_mod)
plot(allEffects(f_fledg_raising_glm),type='response')
```

### GLMM
```{r}
f_fledg_raising_glmm <- glmer(formula = nr_f_fledg ~ raising_type + (1|parent_f), 
                              data = f_fledg_mother, 
                              family = "poisson")
summary(f_fledg_raising_glmm) # not significant
```

#### Plot
```{r}
f_fledg_raising_predi <- ggpredict(f_fledg_raising_glmm, terms = "raising_type")
f_fledg_raising_predi
```


```{r}
plot(x = f_fledg_raising_predi, 
     show.title = F, 
     limits = c(0,1)) + 
  labs(x = "Raising type",
       y = "Number of female fledglings per potential mother", 
       tags = "") +
  theme_classic() + 
  theme( 
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size=14),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    plot.margin = unit(c(7.5, 12.5, 5.5, 15.5), "points"),
    axis.title.x = element_text(colour="black", size = 15),
    axis.title.y = element_text(colour="black",size=15, vjust = 5))
```

#### Testing the models
We use the DHARMa package.

Prepare it
```{r}
f_fledg_raising_simu <- simulateResiduals(fittedModel = f_fledg_raising_glmm)
plot(f_fledg_raising_simu)
```

Dispersion
```{r}
testDispersion(f_fledg_raising_simu)
```

Zero-inflation
```{r}
testZeroInflation(f_fledg_raising_simu)
```

Temporal Autocorrelation
```{r}
testTemporalAutocorrelation(f_fledg_raising_simu)
```


## GLMM of the number of fledglings per potential mother per colony
### Data exploration
```{r}
# Plot the number of fledglings
ggplot(data = f_fledg_mother, mapping = aes(x = nr_f_fledg)) + 
  geom_histogram()

# Plot of the number of fledglings per colony per year
f_fledg_mother %>%
  group_by(breeding_area, hatch_year) %>%
  mutate(sum.fledg = sum(nr_f_fledg)) %>%
  ggplot(mapping = aes(x = hatch_year, y = sum.fledg)) + 
  geom_line(aes(group = breeding_area, color = breeding_area))

# Sum of the breeding events per colony
table(f_fledg_mother$breeding_area)
```

Remove individuals with colony "allocation"
```{r}
f_fledg_mother <- f_fledg_mother[f_fledg_mother$breeding_area != "allocation", ]
f_fledg_mother <- droplevels(f_fledg_mother)
table(f_fledg_mother$breeding_area) # half half
```

### GLM
```{r}
null_model_f_fledg_col <- glm(formula = nr_f_fledg ~ 1, 
                              data = f_fledg_mother, 
                              family = "poisson")
summary(null_model_f_fledg_col)

f_fledg_colony_glm <- glm(formula = nr_f_fledg ~ breeding_area,
                          data = f_fledg_mother,
                          family = "poisson")
summary(f_fledg_colony_glm) # not significant
```

#### Plot
```{r}
#dredge(raising_mod)
plot(allEffects(f_fledg_colony_glm),type='response')
```


### GLMM
```{r}
f_fledg_colony_glmm <- glmer(formula = nr_f_fledg ~ breeding_area + (1|parent_f), 
                             data = f_fledg_mother,  
                             family = "poisson")
summary(f_fledg_colony_glmm)
```

#### Plot
```{r}
f_fledg_colony_predi <- ggpredict(f_fledg_colony_glmm, terms = "breeding_area")
f_fledg_colony_predi
```


```{r}
plot(x = f_fledg_colony_predi, 
     show.title = F, 
     limits = c(0,1.15)) + 
  labs(x = "Colony",
       y = "Number of female fledglings per potential mother", 
       tags = "") +
  theme_classic() + 
  theme( 
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size=14),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    plot.margin = unit(c(7.5, 12.5, 5.5, 15.5), "points"),
    axis.title.x = element_text(colour="black", size = 15),
    axis.title.y = element_text(colour="black",size=15, vjust = 5))
```

#### Testing the models
We use the DHARMa package.

Prepare it
```{r}
f_fledg_colony_simu <- simulateResiduals(fittedModel = f_fledg_colony_glmm)
plot(f_fledg_colony_simu)
```

Dispersion
```{r}
testDispersion(f_fledg_colony_simu)
```

Zero-inflation
```{r}
testZeroInflation(f_fledg_colony_simu)
```

Temporal Autocorrelation
```{r}
testTemporalAutocorrelation(f_fledg_colony_simu)
```

# Female and male fledglings per nest per year
Based on the values for RRNest
## GLMM of the number of fledglings per nest per raising type
```{r}
names(nest_numbers_fm)
```

## Data exploration
```{r}
# Plot the number of fledglings
ggplot(data = nest_numbers_fm, mapping = aes(x = b_count_fledgling)) + 
  geom_histogram()

# Plot of the number of fledglings per raising type per year
nest_numbers_fm %>% 
  group_by(raising_type, breed_year) %>%
  mutate(sum.fledg = sum(b_count_fledgling)) %>%
  ggplot(mapping = aes(x = breed_year, y = sum.fledg)) + 
  geom_line(aes(group = raising_type, color = raising_type))

# Sum of the breeding events with reproduction per nest and per raising type
table(nest_numbers_fm$raising_type)
```

## Analysis with temporarily added females as mothers
### GLM
```{r}
null_model_fm_nest_raising_full <- glm(formula = b_count_fledgling ~ 1, 
                                       data = nest_numbers_fm, 
                                       family = "poisson")
summary(null_model_fm_nest_raising_full)

fm_nest_raising_full_glm <- glm(formula = b_count_fledgling ~ raising_type,
                                data = nest_numbers_fm,
                                family = "poisson")
summary(fm_nest_raising_full_glm) # non significant
```

#### Plot
```{r}
#dredge(raising_mod)
plot(allEffects(fm_nest_raising_full_glm), type='response')
```

### GLMM
```{r}
fm_nest_raising_full_glmm <- glmer(formula = b_count_fledgling ~ raising_type + (1|b_parent_f), 
                                   data = nest_numbers_fm, 
                                   family = "poisson")
summary(fm_nest_raising_full_glmm) # not significant
```

#### Plot
```{r}
fm_nest_raising_full_predi <- ggpredict(fm_nest_raising_full_glmm, terms = "raising_type")
fm_nest_raising_full_predi
```


```{r}
plot(x = fm_nest_raising_full_predi, 
     show.title = F, 
     limits = c(0,6)) + 
  labs(x = "Raising type",
       y = "Number of fledglings per nest (m + f)", 
       tags = "") +
  theme_classic() + 
  theme( 
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size=14),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    plot.margin = unit(c(7.5, 12.5, 5.5, 15.5), "points"),
    axis.title.x = element_text(colour="black", size = 15),
    axis.title.y = element_text(colour="black",size=15, vjust = 5))
```

#### Testing the models
We use the DHARMa package.

Prepare it
```{r}
fm_nest_raising_full_simu <- simulateResiduals(fittedModel = fm_nest_raising_full_glmm)
plot(fm_nest_raising_full_simu)
```

Dispersion
```{r}
testDispersion(fm_nest_raising_full_simu)
```

Zero-inflation
```{r}
testZeroInflation(fm_nest_raising_full_simu)
```

Temporal Autocorrelation
```{r}
testTemporalAutocorrelation(fm_nest_raising_full_simu)
```


## Analysis without temporarily added females as mothers
Here we consider only the raising types foster parent and biological parent

```{r}
nest_numbers_fm_bpfp <- nest_numbers_fm[nest_numbers_fm$raising_type != "added",]
```


### GLM
```{r}
null_model_fm_nest_raising <- glm(formula = b_count_fledgling ~ 1, 
                                  data = nest_numbers_fm_bpfp, 
                                  family = "poisson")
summary(null_model_fm_nest_raising)

fm_nest_raising_glm <- glm(formula = b_count_fledgling ~ raising_type,
                           data = nest_numbers_fm_bpfp,
                           family = "poisson")
summary(fm_nest_raising_glm) # not significant
```

#### Plot
```{r}
#dredge(raising_mod)
plot(allEffects(fm_nest_raising_glm), type='response')
```

### GLMM
```{r}
fm_nest_raising_glmm <- glmer(formula = b_count_fledgling ~ raising_type + (1|b_parent_f), 
                                   data = nest_numbers_fm_bpfp, 
                                   family = "poisson")
summary(fm_nest_raising_glmm) # not significant
```

#### Plot
```{r}
fm_nest_raising_predi <- ggpredict(fm_nest_raising_glmm, terms = "raising_type")
fm_nest_raising_predi
```


```{r}
plot(x = fm_nest_raising_predi, 
     show.title = F, 
     limits = c(0,6)) + 
  labs(x = "Raising type",
       y = "Number of fledglings per nest (m + f)", 
       tags = "") +
  theme_classic() + 
  theme( 
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size=14),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    plot.margin = unit(c(7.5, 12.5, 5.5, 15.5), "points"),
    axis.title.x = element_text(colour="black", size = 15),
    axis.title.y = element_text(colour="black",size=15, vjust = 5))
```

#### Testing the models
Testing with the DHARMa package.

Prepare it
```{r}
fm_nest_raising_simu <- simulateResiduals(fittedModel = fm_nest_raising_glmm)
plot(fm_nest_raising_full_simu)
```

Dispersion
```{r}
testDispersion(fm_nest_raising_simu)
```

Zero-inflation
```{r}
testZeroInflation(fm_nest_raising_simu)
```

Temporal Autocorrelation
```{r}
testTemporalAutocorrelation(fm_nest_raising_simu)
```


# GLMM of the number of fledglings per nest per colony

## Data exploration
```{r}
# Plot the number of fledglings
ggplot(data = nest_numbers_fm, mapping = aes(x = b_count_fledgling)) + 
  geom_histogram()

# Plot of the number of fledglings per raising type per year
nest_numbers_fm %>% 
  group_by(b_breeding_colony, breed_year) %>%
  mutate(sum.fledg = sum(b_count_fledgling)) %>%
  ggplot(mapping = aes(x = breed_year, y = sum.fledg)) + 
  geom_line(aes(group = b_breeding_colony, color = b_breeding_colony))

# Sum of the breeding events with reproduction per nest and per raising type
table(nest_numbers_fm$b_breeding_colony)
```

## GLM
```{r}
null_model_fm_nest_colony <- glm(formula = b_count_fledgling ~ 1, 
                                 data = nest_numbers_fm, 
                                 family = "poisson")
summary(null_model_fm_nest_colony)

fm_nest_colony_glm <- glm(formula = b_count_fledgling ~ b_breeding_colony,
                          data = nest_numbers_fm,
                          family = "poisson")
summary(fm_nest_colony_glm) # non significant
```

### Plot
```{r}
#dredge(raising_mod)
plot(allEffects(fm_nest_colony_glm), type='response')
```

## GLMM
```{r}
fm_nest_colony_glmm <- glmer(formula = b_count_fledgling ~ b_breeding_colony + (1|b_parent_f), 
                             data = nest_numbers_fm, 
                             family = "poisson")
summary(fm_nest_colony_glmm) # not significant
```

### Plot
```{r}
fm_nest_colony_predi <- ggpredict(fm_nest_colony_glmm, terms = "b_breeding_colony")
fm_nest_colony_predi
```


```{r}
plot(x = fm_nest_colony_predi, 
     show.title = F, 
     limits = c(0,3)) + 
  labs(x = "Breeding colony",
       y = "Number of fledglings per nest (m + f)", 
       tags = "") +
  theme_classic() + 
  theme( 
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size=14),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    plot.margin = unit(c(7.5, 12.5, 5.5, 15.5), "points"),
    axis.title.x = element_text(colour="black", size = 15),
    axis.title.y = element_text(colour="black",size=15, vjust = 5))
```

### Testing the models
Testing with the DHARMa package.

Prepare it
```{r}
fm_nest_colony_simu <- simulateResiduals(fittedModel = fm_nest_colony_glmm)
plot(fm_nest_colony_simu)
```

Dispersion
```{r}
testDispersion(fm_nest_colony_simu)
```

Zero-inflation
```{r}
testZeroInflation(fm_nest_colony_simu)
```

Temporal Autocorrelation
```{r}
testTemporalAutocorrelation(fm_nest_colony_simu)
```


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

