---
title: "PVA of the NBI: Analysis of Management Improvement Scenarios - Data Preparation" ## name of your project
author: "Sinah Drenske, Cédric Scherer"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 9, fig.height = 6, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Hypothesis:** We hypothesize that at the present time the northern bald ibis population can survive without further management and release. We predict that the observed demographic rates will ensure population growth and do not differ between the breeding colonies. 
* **Study area:** Austria, Germany and Italy
* **Data:** Results of the first NetLogo Model

In this script we prepare the data of the first NetLogo model for the analysis.

# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("here")          ## loading data
library("tidyverse")     ## data manipulation
library("magrittr")      ## Pipes
library("data.table")    ## fread() and fwrite()
```


# Data
Load data tables, increase run numbers of the other partial models and combine them. We had to split up the model in 4 partial models, because of different model options (df_Basic, df_Baseline_SD, df_All_chicks) or because of different model running time (50 years for the model vs. 100 years for a simulation of the baseline scenario).
df Basic is a model with all possible combinations of the baseline survival and reproduction values and their improvements.
df_Baseline_SD is a model where we included the standard deviation in the NetLogo model. We did not do that for the other models.
df_All_chicks is a model where we combined the reproduction values "Status quo" or "All chicks" only with the Baseline survival probabilities. 
df_Baseline_100 is a model where the running time was for 100 years not 50 years as for all the other models.

```{r data}
# Baseline without SD and improvements of survival and reproduction
df_Basic <- 
  df_Basic_raw <- fread(file = here("data-raw", "model-raw", "NBI_Baseline_and_Improvements.csv"),
                        skip = 6)

# Baseline with SD
df_Baseline_SD <- 
  df_Baseline_SD_raw <- fread(file = here("data-raw", "model-raw", "NBI_Baseline_SD.csv"),
                              skip = 6)

## Increase the run number so we have each run number only once
df_Baseline_SD$`[run number]` <- as.numeric(as.character(df_Baseline_SD$`[run number]`))
df_Baseline_SD$`[run number]` <- df_Baseline_SD$`[run number]` + 50000


# Status quo & All chicks
df_All_Chicks <- 
  df_All_Chicks_raw <- fread(file = here("data-raw", "model-raw", "NBI_All_chicks_Statquo.csv"), 
                             skip = 6)

## Increase the run number so we have each run number only once
df_All_Chicks$`[run number]` <- as.numeric(as.character(df_All_Chicks$`[run number]`))
df_All_Chicks$`[run number]` <- df_All_Chicks$`[run number]` + 60000

# Baseline for 100 years
df_Baseline_100 <- 
  df_Baseline_SD_raw <- fread(file = here("data-raw", "model-raw", "NBI_Baseline_SD_100years.csv"),
                              skip = 6)
```

Combine the datasets 
```{r}
df_combine <- rbind(df_Baseline_SD, df_All_Chicks)
df_combine$`[run number]` <- as.character(df_combine$`[run number]`)
table(df_combine$`[run number]` < 60000) # 6956 Rows from Baseline_SD, increased by 50 000 not 60 000 
```

Delete Baseline scenario without SD in df_Basic
```{r}
# The description of the Baseline scenario can be found in the respective column of the table of the baseline simulation with SD
unique(df_Baseline_SD$`(word Mortality_Juveniles Mortality_Subadults_Age1 Mortality_Subadults_Age2 Mortality_Adults Fecundity Supplements? Number_Supplements Supplement_Time Catastrophes? Frequency Severity)`)
```

```{r}
# How many rows will be removed?
table(df_Basic$`(word Mortality_Juveniles Mortality_Subadults_Age1 Mortality_Subadults_Age2 Mortality_Adults Fecundity Supplements? Number_Supplements Supplement_Time Catastrophes? Frequency Severity)` == "0.360.260.310.220.53true04true00.25")

# Remove the baseline scenario without SD
df_Basic <- df_Basic[df_Basic$`(word Mortality_Juveniles Mortality_Subadults_Age1 Mortality_Subadults_Age2 Mortality_Adults Fecundity Supplements? Number_Supplements Supplement_Time Catastrophes? Frequency Severity)` != "0.360.260.310.220.53true04true00.25"]
```

Combine df_Basic and df_combine
```{r}
df_Basic <- rbind(df_Basic, df_combine)
```


# Data preparation
```{r}
head(df_Basic)
```

## Rename columns
```{r}
names(df_Basic)
```

```{r}
rename_map <- c(
  "[run number]"                                                                      = "Run",
  "Number_Juveniles"                                                                  = "N_Juveniles",
  "Number_Subadults_Age1"                                                             = "N_Subadults1",
  "Number_Subadults_Age2"                                                             = "N_Subadults2",
  "Number_Adults"                                                                     = "N_Adults",
  "Mortality_Juveniles"                                                               = "Mortality_Juveniles",
  "Mortality_Subadults_Age1"                                                          = "Mortality_Subadults1",
  "Mortality_Subadults_Age2"                                                          = "Mortality_Subadults2",
  "Mortality_Adults"                                                                  = "Mortality_Adults",
  "Fecundity"                                                                         = "Repro",
  "Supplements?"                                                                      = "Supplemented",
  "Number_Supplements"                                                                = "N_Supplements",
  "Supplement_Time"                                                                   = "Supplement_Time",
  "Catastrophes?"                                                                     = "Stoch_events",
  "Frequency"                                                                         = "Stoch_Frequency",
  "Severity"                                                                          = "Stoch_Severity",
  "[step]"                                                                            = "Year",
  "(word Mortality_Juveniles Mortality_Subadults_Age1 Mortality_Subadults_Age2 Mortality_Adults Fecundity Supplements? Number_Supplements Supplement_Time Catastrophes? Frequency Severity)" = "Scenario",
  "count turtles"                                                                     = "IndAll",
  "count turtles with [colony = \"\"Burghausen\"\"]"                                  = "IndB",
  "count turtles with [colony = \"\"Kuchl\"\"]"                                       = "IndK",
  "count turtles with [colony = \"\"Ueberlingen\"\"]"                                 = "IndUb",
  "count turtles with [raising = \"\"FP\"\"]"                                         = "IndFP",
  "count turtles with [raising = \"\"P\"\"]"                                          = "IndP",
  "count turtles with [colony = \"\"Burghausen\"\" and raising = \"\"FP\"\"]"         = "IndBFP",
  "count turtles with [colony = \"\"Kuchl\"\" and raising = \"\"FP\"\"]"              = "IndKFP",
  "count turtles with [colony = \"\"Ueberlingen\"\" and raising = \"\"FP\"\"]"        = "IndUbFP",
  "count turtles with [colony = \"\"Burghausen\"\" and raising = \"\"P\"\"]"          = "IndBP",
  "count turtles with [colony = \"\"Kuchl\"\" and raising = \"\"P\"\"]"               = "IndKP",
  "count turtles with [colony = \"\"Ueberlingen\"\" and raising = \"\"P\"\"]"         = "IndUbP",
  "count turtles with [age = 0]"                                                      = "IndJuv",
  "count turtles with [age = 1]"                                                      = "IndSub1",
  "count turtles with [age = 2]"                                                      = "IndSub2",
  "count turtles with [age >= 3]"                                                     = "IndAdu")

names(df_Basic) <- rename_map[names(df_Basic)]
names(df_Baseline_100) <- rename_map[names(df_Baseline_100)]
rm(rename_map)
head(df_Basic)
```

## Factorize & numerize variables
```{r}
str(df_Basic)
df_Basic %<>% 
  mutate(Scenario     = factor(Scenario),
         Supplemented = factor(Supplemented),
         Stoch_events = factor(Stoch_events))
df_Basic %<>%
  mutate(Run                    = as.numeric(as.character(Run)),
         N_Juveniles            = as.numeric(as.character(N_Juveniles)),
         N_Subadults1           = as.numeric(as.character(N_Subadults1)),
         N_Subadults2           = as.numeric(as.character(N_Subadults2)),
         N_Adults               = as.numeric(as.character(N_Adults)),
         Mortality_Juveniles    = as.numeric(as.character(Mortality_Juveniles)),
         Mortality_Subadults1   = as.numeric(as.character(Mortality_Subadults1)),
         Mortality_Subadults2   = as.numeric(as.character(Mortality_Subadults2)),
         Mortality_Adults       = as.numeric(as.character(Mortality_Adults)),
         Repro                  = as.numeric(as.character(Repro)),
         N_Supplements          = as.numeric(as.character(N_Supplements)),
         Supplement_Time        = as.numeric(as.character(Supplement_Time)),
         Stoch_Frequency        = as.numeric(as.character(Stoch_Frequency)),
         Stoch_Severity         = as.numeric(as.character(Stoch_Severity)),
         Year                   = as.numeric(as.character(Year)),
         IndAll                 = as.numeric(as.character(IndAll)),
         IndB                   = as.numeric(as.character(IndB)),
         IndK                   = as.numeric(as.character(IndK)),
         IndUb                  = as.numeric(as.character(IndUb)),
         IndFP                  = as.numeric(as.character(IndFP)),
         IndP                   = as.numeric(as.character(IndP)),
         IndBFP                 = as.numeric(as.character(IndBFP)),
         IndKFP                 = as.numeric(as.character(IndKFP)),
         IndUbFP                = as.numeric(as.character(IndUbFP)),
         IndBP                  = as.numeric(as.character(IndBP)),
         IndKP                  = as.numeric(as.character(IndKP)),
         IndUbP                 = as.numeric(as.character(IndUbP)),
         IndJuv                 = as.numeric(as.character(IndJuv)),
         IndSub1                = as.numeric(as.character(IndSub1)),
         IndSub2                = as.numeric(as.character(IndSub2)),
         IndAdu                 = as.numeric(as.character(IndAdu)))

str(df_Basic)
```

```{r}
# DO the same for the simulation for 100 years
df_Baseline_100 %<>% 
  mutate(Scenario     = factor(Scenario),
         Supplemented = factor(Supplemented),
         Stoch_events = factor(Stoch_events))
df_Baseline_100 %<>%
  mutate(Run                    = as.numeric(as.character(Run)),
         N_Juveniles            = as.numeric(as.character(N_Juveniles)),
         N_Subadults1           = as.numeric(as.character(N_Subadults1)),
         N_Subadults2           = as.numeric(as.character(N_Subadults2)),
         N_Adults               = as.numeric(as.character(N_Adults)),
         Mortality_Juveniles    = as.numeric(as.character(Mortality_Juveniles)),
         Mortality_Subadults1   = as.numeric(as.character(Mortality_Subadults1)),
         Mortality_Subadults2   = as.numeric(as.character(Mortality_Subadults2)),
         Mortality_Adults       = as.numeric(as.character(Mortality_Adults)),
         Repro                  = as.numeric(as.character(Repro)),
         N_Supplements          = as.numeric(as.character(N_Supplements)),
         Supplement_Time        = as.numeric(as.character(Supplement_Time)),
         Stoch_Frequency        = as.numeric(as.character(Stoch_Frequency)),
         Stoch_Severity         = as.numeric(as.character(Stoch_Severity)),
         Year                   = as.numeric(as.character(Year)),
         IndAll                 = as.numeric(as.character(IndAll)),
         IndB                   = as.numeric(as.character(IndB)),
         IndK                   = as.numeric(as.character(IndK)),
         IndUb                  = as.numeric(as.character(IndUb)),
         IndFP                  = as.numeric(as.character(IndFP)),
         IndP                   = as.numeric(as.character(IndP)),
         IndBFP                 = as.numeric(as.character(IndBFP)),
         IndKFP                 = as.numeric(as.character(IndKFP)),
         IndUbFP                = as.numeric(as.character(IndUbFP)),
         IndBP                  = as.numeric(as.character(IndBP)),
         IndKP                  = as.numeric(as.character(IndKP)),
         IndUbP                 = as.numeric(as.character(IndUbP)),
         IndJuv                 = as.numeric(as.character(IndJuv)),
         IndSub1                = as.numeric(as.character(IndSub1)),
         IndSub2                = as.numeric(as.character(IndSub2)),
         IndAdu                 = as.numeric(as.character(IndAdu)))
```


## Reorder rows
```{r}
df_Basic  %<>%
  arrange(Run, Year)%>%
  tibble::as_tibble()

# for 100 years
df_Baseline_100  %<>%
  arrange(Run, Year)%>%
  tibble::as_tibble()
```

## Reorder columns
```{r}
df_Basic %<>% 
  select(Run, Scenario, Year, 
         IndAll,IndP,
         IndJuv,IndSub1,IndSub2,IndAdu,
         Mortality_Juveniles,Mortality_Subadults1,Mortality_Subadults2,Mortality_Adults,Repro,
         Supplemented,N_Supplements,Supplement_Time, Stoch_events, Stoch_Frequency, Stoch_Severity,
         N_Juveniles,N_Subadults1,N_Subadults2,N_Adults)

# For 100 years
df_Baseline_100 %<>% 
  select(Run, Scenario, Year, 
         IndAll,IndP,
         IndJuv,IndSub1,IndSub2,IndAdu,
         Mortality_Juveniles,Mortality_Subadults1,Mortality_Subadults2,Mortality_Adults,Repro,
         Supplemented,N_Supplements,Supplement_Time, Stoch_events, Stoch_Frequency, Stoch_Severity,
         N_Juveniles,N_Subadults1,N_Subadults2,N_Adults)
```


# Save data
```{r}
# saveRDS(object = df_Basic, file = here("output", "data-proc", "model-proc", "df_Basic.rds"))
# saveRDS(object = df_Baseline_100, file = here("output", "data-proc", "model-proc", "df_Baseline_100.rds"))
```


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

