---
title: "PVA of the NBI: Analysis of Management Improvement Scenarios - Analysis" ## name of your project
author: "Your Name"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 9, fig.height = 6, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Research question:** (Q1) How will the NBI population develop in the next 50 years and what will be the extinction risk, given that the colonies will be either managed or left to themselves? (Q2) Will the developments in the two already established colonies Burghausen & Kuchl be different? 
* **Study area:** Austria, Germany and Italy
* **Data:** Results of the first NetLogo Model


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("here")  
library("tidyverse")     ## data manipulation
library("magrittr")
```


# Data

```{r data}
df_Basic <- readRDS(file = here("output", "data-proc", "NetLogo-proc", "df_Basic.rds"))

df_Basic_100 <- readRDS(file = here("output", "data-proc", "NetLogo-proc", "df_Baseline_100.rds"))
```


# Data overview
## Minimum
```{r}
df_min <- df_Basic %>% group_by(Run, Scenario) %>% summarize_if(is.numeric, min) # lowest value per run
df_min_100 <- df_Basic_100 %>% group_by(Run, Scenario) %>% summarize_if(is.numeric, min)
```

```{r}
range(df_min$IndAll) # range of the lowest values
range(df_min_100$IndAll)
```

```{r}
df_min <- droplevels(df_min)
df_min_100 <- droplevels(df_min_100)
```

```{r}
# How often went the population extinct per scenario (100 runs)? TRUE = went extinct, FALSE = went not extinct
df_min_tab <- table(df_min$Scenario, df_min$IndAll < 1) 
table(df_min$IndAll < 1)

df_min_tab_100 <- table(df_min_100$Scenario, df_min_100$IndAll < 1) 
table(df_min_100$IndAll < 1)
```

```{r}
# Have a look at the table
# There is one row per scenario with the numbers of scenarios that went extinct or not
head(df_min_tab)
tail(df_min_tab)
```

```{r}
# Baseline 100 years
df_min_tab_100
```

## Extinctions
```{r}
ext <- df_min_tab[,2]
ext_100 <- df_min_tab_100[,2]
```

```{r}
df_ext <- df_min %>% group_by(Scenario) %>% summarize_if(is.numeric, min) # calculate the lowest value of all runs per scenario
df_ext$Extinct <- ext

df_ext_100 <- df_min_100 %>% group_by(Scenario) %>% summarize_if(is.numeric, min) # calculate the lowest value of all runs per scenario
df_ext_100$Extinct <- ext_100
```

```{r}
table(df_ext$Extinct) # Table the number of runs that went extinct. In 316 scenarios the population never went extinct
table(df_ext_100$Extinct) # in 87 of 100 repetitions (87%) the population went extinct when model ran for 100 years
```

```{r}
table(df_ext$Extinct < 6) # Extinction probability: How often went the population extinct 5 times or less?
# In 324 of 326 scenarios the population went not extinct more than 5 times of 100 runs (Extinction probability <= 5%)
```

```{r}
rm(df_min_tab, ext)
rm(df_min_tab_100, ext_100)
```

## Maximum
```{r}
df_max <- df_Basic %>% group_by(Run,Scenario) %>% summarize_if(is.numeric, max) # highest value per run
```

```{r}
range(df_max$IndAll) # 74 - 7146
```

```{r}
df_max <- droplevels(df_max)
df_max_sc <- df_max %>% group_by(Scenario) %>% summarize_if(is.numeric, max) # highest value per scenario
```

## Save data
```{r}
# saveRDS(object = df_ext, file = here("output", "data-proc", "NetLogo-proc", "df_ext.rds"))
# saveRDS(object = df_max_sc, file = here("output", "data-proc", "NetLogo-proc", "df_max_sc.rds"))
```


# Calculation of Lambda
Calculate the difference between the years per run, put NA for the first value of each run (because there you can not calculate a difference)
```{r}
df_Basic <- df_Basic %>%  group_by(Run) %>% mutate(Diff = c(NA,diff(IndAll)))
```

Calculate the ratio (quotient) between the difference (step before) and the number of individuals in the year before
```{r}
df_Basic <- df_Basic %>%  group_by(Run) %>% mutate(Quot = Diff / lag(IndAll,1))
```

Calculate the stochastic Lambda by add 1 to the ratio (step before). Now we have a Lambda per year per run
```{r}
df_Basic <- df_Basic %>%  group_by(Run) %>% mutate(Lambda_sto = Quot + 1)
```

```{r}
range(df_Basic$Lambda_sto, na.rm = T) # 0 2
df_Basic <- df_Basic[, -c(35,36)] # remove the first columns for the Lambda calculation
```



## Other possibility to calculate Lambda
```{r}
#df_Basic <- df_Basic %>%  group_by(Run) %>% mutate(Lambda2 = IndAll/ lag(IndAll,1))
```

```{r}
# # Analyse the differences
# table(round(df_Basic$Lambda_sto, digits = 5) == round(df_Basic$Lambda2, digits = 5), useNA = "always")
# which(round(df_Basic$Lambda_sto, digits = 5) != round(df_Basic$Lambda2, digits = 5))
# 
# table(round(df_Basic$Lambda_sto, digits = 2) == round(df_Basic$Lambda2, digits = 2), useNA = "always")

```


## Save data
Save a new version of df_Basic with Lambda
```{r}
#saveRDS(object = df_Basic, file = here("output", "data-proc", "NetLogo-proc", "df_Basic_Lambda.rds"))
```


# Mean per scenario
At first we calculate mean values per run. E.g. there is one value per run for Lambda or for the number of individuals during this run for all years of this run.
```{r}
# Mean for each run
df_run <- df_Basic %>%  group_by(Run, Scenario) %>% summarize_if(is.numeric, mean, na.rm = TRUE)
# 32600 rows, 1 row for each Run. This is needed to calculate the SD of Lambda
```

Calculate the SD of the stochastic Lambda per scenario
```{r}
df_scenario <- df_run %>% group_by(Scenario) %>% mutate(SD = sd(Lambda_sto, na.rm = T))
```

Calculate the SD
```{r}
df_sc_sd <- df_scenario %>% group_by(Scenario) %>% summarize(SD = mean(SD, na.rm = T))
```

Now there will be one value for each column per scenario
```{r}
df_sc_mean <- df_Basic %>% group_by(Scenario) %>% summarize_if(is.numeric, mean, na.rm = TRUE)
```

Combine the SD and Extinction with df_Basic
```{r}
df_sc_mean$SD_Lambda <- df_sc_sd$SD
df_sc_mean$Extinct <- df_ext$Extinct
rm(df_sc_sd)
```


## Overview of the data
```{r}
range(df_sc_mean$Lambda_sto)
# 0.9538676 1.3987626
table(df_sc_mean$Lambda_sto > 1) 
# TRUE for 308 scenarios - positive population growth in 308 scenarios

table(df_sc_mean$Extinct <= 5) 
# Extinction probability 5% or less in 324 scenarios

table(df_sc_mean$Lambda_sto > 1 & df_sc_mean$Extinct <= 5) 
# Positive population growth and an extinction probability of 5% or less in 308 scenarios of 326 (94.48%)
```

## Save data 
```{r}
#saveRDS(object = df_sc_mean, file = here("output", "data-proc", "NetLogo-proc", "df_sc_mean.rds"))
```


# Analysis I: Scenarios with positive population growth
```{r}
df_lambda_pos <- subset(df_sc_mean, df_sc_mean$Lambda_sto > 1 & df_sc_mean$Extinct <= 5) # 308 rows
df_lambda_pos <- droplevels(df_lambda_pos)
```

Overview
```{r}
range(df_lambda_pos$Lambda_sto)
table(df_lambda_pos$Extinct)
```

Table the distributions between the different mortalities and reproduction rates
```{r}
table(df_lambda_pos$Mortality_Juveniles)
# 0.20  0.30  0.36 
#  107   102    99          
table(df_lambda_pos$Mortality_Subadults1)
# 0.08  0.19  0.26   
#  107   102    99          
table(df_lambda_pos$Mortality_Subadults2)
# 0.14  0.24  0.31 
#  107   102    99          
table(df_lambda_pos$Mortality_Adults)
# 0.02  0.14  0.22 
#  108   108    92          
table(df_lambda_pos$Repro)
# 0.53  0.58  0.66  1.06  1.41  3.97 
#   71    74    80    81     1     1   
```


## Save data 
```{r}
#saveRDS(object = df_lambda_pos, file = here("output", "data-proc", "NetLogo-proc", "df_lambda_pos.rds"))
```

# Analysis II: Scenarios of special interest
Find the rows of specific scenarios
```{r}
# Baseline
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 0.53)


# Better reproduction
## Reproduction increased by 10%
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 0.58)

## Reproduction increased by 25%
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 0.66)

## Reproduction increased by 100%
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 1.06)


# Better juveniles survival (s1)
## s1 increased by 10%
which(df_sc_mean$Mortality_Juveniles == 0.3 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 0.53)

## s1 increased by 25%
which(df_sc_mean$Mortality_Juveniles == 0.2 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 0.53)


# Better adult survival (s4)
## s4 increased by 10%
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.14 & df_sc_mean$Repro == 0.53)

## s4 increased by 25%
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.02 & df_sc_mean$Repro == 0.53)

# Better survival all stages (s1-s4)
## s1-s4 increased by 10%
which(df_sc_mean$Mortality_Juveniles == 0.3 & df_sc_mean$Mortality_Subadults1 == 0.19 & df_sc_mean$Mortality_Subadults2 == 0.24 & df_sc_mean$Mortality_Adults == 0.14 & df_sc_mean$Repro == 0.53)

## s1-s4 increased by 25%
which(df_sc_mean$Mortality_Juveniles == 0.2 & df_sc_mean$Mortality_Subadults1 == 0.08 & df_sc_mean$Mortality_Subadults2 == 0.14 & df_sc_mean$Mortality_Adults == 0.02 & df_sc_mean$Repro == 0.53)


# Better survival all stages & better reproduction (s1-s4 + reproduction)
## s1-s4 + reproduction increased by 10%
which(df_sc_mean$Mortality_Juveniles == 0.3 & df_sc_mean$Mortality_Subadults1 == 0.19 & df_sc_mean$Mortality_Subadults2 == 0.24 & df_sc_mean$Mortality_Adults == 0.14 & df_sc_mean$Repro == 0.58)

## s1-s4 + reproduction increased by 25%
which(df_sc_mean$Mortality_Juveniles == 0.2 & df_sc_mean$Mortality_Subadults1 == 0.08 & df_sc_mean$Mortality_Subadults2 == 0.14 & df_sc_mean$Mortality_Adults == 0.02 & df_sc_mean$Repro == 0.66)


# Status quo 
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 1.41)


# All chicks
which(df_sc_mean$Mortality_Juveniles == 0.36 & df_sc_mean$Mortality_Subadults1 == 0.26 & df_sc_mean$Mortality_Subadults2 == 0.31 & df_sc_mean$Mortality_Adults == 0.22 & df_sc_mean$Repro == 3.97)

```

Build a table with the scenarios
```{r}
df_scen <- df_sc_mean[c( 321,322,323,324, 213,105,317,313, 161,1,162,3, 325,326),] 
df_scen$Scenario_name <- c("Baseline","10RR", "25RR","100RR",
                           "10juv","25juv","10adu","25adu",
                           "10all","25all","10all10RR","25all25RR",
                           "Statusquo", "All_chicks")
```

delete & reorder columns
```{r}
df_scen <- df_scen[,-c(2,3,25:32)] # Run, Year, Catastrophe settings (not used yet), settings for the starting population
df_scen <- droplevels(df_scen)

# Which columns are still there?
names(df_scen)
```

Reorder the columns
```{r}
df_scen %<>%
  select(Scenario_name,Scenario, Lambda_sto,SD_Lambda,Extinct,
         IndAll,IndB,IndK,IndUb,IndFP,IndP,                
         IndBFP,IndBP,IndKFP,IndKP,IndUbFP,IndUbP,
         IndJuv,IndSub1,IndSub2,IndAdu,         
         Mortality_Juveniles,Mortality_Subadults1,Mortality_Subadults2,Mortality_Adults,Repro)
```

## Save data
```{r}
#saveRDS(object = df_scen, file = here("output", "data-proc", "NetLogo-proc", "df_scen.rds"))
```


## Scenarios where Lambda > 1
```{r}
df_scen_Lambda <- df_scen[which(df_scen$Lambda_sto > 1 & df_scen$Extinct < 6), ] # 9 rows
```

## Calculation of the mean per year per Scenario
```{r}
levels(df_scen$Scenario)
```

Subset all rows from df_Basic of the 14 scenarios of interest
```{r}
df_Basic_scen <- subset(df_Basic,
                        df_Basic$Scenario == "0.20.080.140.020.53true04true00.25" |
                        df_Basic$Scenario == "0.20.080.140.020.66true04true00.25" |
                        df_Basic$Scenario == "0.20.260.310.220.53true04true00.25" |
                        df_Basic$Scenario == "0.30.190.240.140.53true04true00.25" |
                        df_Basic$Scenario == "0.30.190.240.140.58true04true00.25" | # 5
                          
                        df_Basic$Scenario == "0.30.260.310.220.53true04true00.25" |
                        df_Basic$Scenario == "0.360.260.310.020.53true04true00.25" |
                        df_Basic$Scenario == "0.360.260.310.140.53true04true00.25" |
                        df_Basic$Scenario == "0.360.260.310.220.53true04true00.25" |
                        df_Basic$Scenario == "0.360.260.310.220.58true04true00.25" | # 10
                          
                        df_Basic$Scenario == "0.360.260.310.220.66true04true00.25" |
                        df_Basic$Scenario == "0.360.260.310.221.06true04true00.25" |
                        df_Basic$Scenario == "0.360.260.310.221.41true04true00.25" |
                        df_Basic$Scenario == "0.360.260.310.223.97true04true00.25")  # 14

df_Basic_scen <- droplevels(df_Basic_scen)
```

```{r}
levels(df_Basic_scen$Scenario) # 14
table(df_Basic_scen$Scenario)
```
### Calculate the mean per year per scenario
```{r}
df_mean <- df_Basic_scen %>% group_by(Scenario,Year) %>% summarize_if(is.numeric, mean)
```

```{r}
table(df_mean$Year)
```

### Save data
```{r}
# saveRDS(object = df_mean, file = here("output", "data-proc", "NetLogo-proc", "df_mean.rds"))
# saveRDS(object = df_Basic_scen, file = here("output", "data-proc", "NetLogo-proc", "df_Basic_scen.rds"))
```


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

