---
title: "PVA of the NBI: Analysis of Stochastic Event Scenarios - Data Preparation" ## name of your project
author: "Sinah Drenske, Cédric Scherer"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 9, fig.height = 6, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Hypothesis:** We hypothesize that at the given stage the northern bald ibis population can survive without further management and release. We predict that observed demographic rates will ensure population growth and do not differ between the colonies 
* **Study area:** Austria, Germany and Italy
* **Data:** Results of the second NetLogo model


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("here")
library("tidyverse")     ## data manipulation
library("magrittr")      ## Pipes
library("data.table")    ## fread() and fwrite()
```


# Data
Load data tables, increase run numbers of the other partial models and combine them. Here we splitted the models because of RAM or a different option for the supplement scenarios. The All Chicks scenario was modelled with and without supplements. 
```{r data}
# Improvements of survival and reproduction
df_StevS <- df_StevS_raw <- fread(
                              file = here("data-raw", "model-raw", "NBI_StevS_Improvements.csv"),
                              skip = 6)

# Baseline with SD 
df_StevS_Baseline <- fread(
                       file = here("data-raw", "model-raw", "NBI_StevS_Baseline_SD.csv"),
                       skip = 6)

df_StevS_Baseline$`[run number]` <- as.numeric(as.character(df_StevS_Baseline$`[run number]`))
df_StevS_Baseline$`[run number]` <- df_StevS_Baseline$`[run number]` + 70000
df_StevS_Baseline$`[run number]` <- as.character(df_StevS_Baseline$`[run number]`)

# All chicks
df_StevS_All_chicks <- fread(
                        file = here("data-raw", "model-raw", "NBI_StevS_All_chicks.csv"), 
                        skip = 6)

df_StevS_All_chicks$`[run number]` <- as.numeric(as.character(df_StevS_All_chicks$`[run number]`))
df_StevS_All_chicks$`[run number]` <- df_StevS_All_chicks$`[run number]` + 80000
df_StevS_All_chicks$`[run number]` <- as.character(df_StevS_All_chicks$`[run number]`)

# All chicks without supplements
df_Stev_All_chicks <- fread(
                        file = here("data-raw", "model-raw", "NBI_Stev_All_chicks_without_Supp.csv"), 
                        skip = 6)

df_Stev_All_chicks$`[run number]` <- as.numeric(as.character(df_Stev_All_chicks$`[run number]`))
df_Stev_All_chicks$`[run number]` <- df_Stev_All_chicks$`[run number]` + 90000
df_Stev_All_chicks$`[run number]` <- as.character(df_Stev_All_chicks$`[run number]`)
```

## Combine the tables
```{r}
df_StevS <- rbind(df_StevS, df_StevS_Baseline, df_StevS_All_chicks, df_Stev_All_chicks)
```

# Data Preparation
```{r}
names(df_StevS)
```

## Rename columns
```{r}
rename_map <- c(
  "[run number]"                                                                      = "Run",
  "scenario"                                                                          = "Over_Scenario",
  "Number_Juveniles"                                                                  = "N_Juveniles",
  "Number_Subadults_Age1"                                                             = "N_Subadults1",
  "Number_Subadults_Age2"                                                             = "N_Subadults2",
  "Number_Adults"                                                                     = "N_Adults",
  "Supplements?"                                                                      = "Supplemented",
  "Number_Supplements"                                                                = "N_Supplements",
  "Supplement_Time"                                                                   = "Supplement_Time",
  "Catastrophes?"                                                                     = "Stoch_events",
  "Frequency"                                                                         = "Stoch_Frequency",
  "Severity"                                                                          = "Stoch_Severity",
  "[step]"                                                                            = "Year",
  "(word scenario Supplements? Number_Supplements Supplement_Time Catastrophes? Frequency Severity)"   = "Scenario_detail",
  "Mortality_Juveniles"                                                               = "Mortality_Juveniles",
  "Mortality_Subadults_Age1"                                                          = "Mortality_Subadults1",
  "Mortality_Subadults_Age2"                                                          = "Mortality_Subadults2",
  "Mortality_Adults"                                                                  = "Mortality_Adults",
  "Fecundity"                                                                         = "Repro",
  "count turtles"                                                                     = "IndAll",
  "count turtles with [colony = \"\"Burghausen\"\"]"                                  = "IndB",
  "count turtles with [colony = \"\"Kuchl\"\"]"                                       = "IndK",
  "count turtles with [colony = \"\"Ueberlingen\"\"]"                                 = "IndUb",
  "count turtles with [raising = \"\"FP\"\"]"                                         = "IndFP",
  "count turtles with [raising = \"\"P\"\"]"                                          = "IndP",
  "count turtles with [colony = \"\"Burghausen\"\" and raising = \"\"FP\"\"]"         = "IndBFP",
  "count turtles with [colony = \"\"Kuchl\"\" and raising = \"\"FP\"\"]"              = "IndKFP",
  "count turtles with [colony = \"\"Ueberlingen\"\" and raising = \"\"FP\"\"]"        = "IndUbFP",
  "count turtles with [colony = \"\"Burghausen\"\" and raising = \"\"P\"\"]"          = "IndBP",
  "count turtles with [colony = \"\"Kuchl\"\" and raising = \"\"P\"\"]"               = "IndKP",
  "count turtles with [colony = \"\"Ueberlingen\"\" and raising = \"\"P\"\"]"         = "IndUbP",
  "count turtles with [age = 0]"                                                      = "IndJuv",
  "count turtles with [age = 1]"                                                      = "IndSub1",
  "count turtles with [age = 2]"                                                      = "IndSub2",
  "count turtles with [age >= 3]"                                                     = "IndAdu")

names(df_StevS) <- rename_map[names(df_StevS)]
rm(rename_map)
head(df_StevS)
```
## Combine Over_Scenario & Scenario detail
```{r}
df_StevS$Scenario <- paste(df_StevS$Over_Scenario, df_StevS$Scenario_detail, sep = "")
unique(df_StevS$Over_Scenario) # 11 (All chicks 2 x, 1 of them without supplements)
# unique(df_StevS$Scenario)
```

Delete column Scenario_detail
```{r}
df_StevS <- df_StevS[, -14] # Delete Scenario_detail
```

## Factorize & numerize variables
```{r}
str(df_StevS)
df_StevS %<>% 
  mutate(Over_Scenario = factor(Over_Scenario),
         Scenario      = factor(Scenario),
         Supplemented  = factor(Supplemented),
         Stoch_events  = factor(Stoch_events))

df_StevS %<>%
  mutate(Run                    = as.numeric(as.character(Run)),
         N_Juveniles            = as.numeric(as.character(N_Juveniles)),
         N_Subadults1           = as.numeric(as.character(N_Subadults1)),
         N_Subadults2           = as.numeric(as.character(N_Subadults2)),
         N_Adults               = as.numeric(as.character(N_Adults)),
         Mortality_Juveniles    = as.numeric(as.character(Mortality_Juveniles)),
         Mortality_Subadults1   = as.numeric(as.character(Mortality_Subadults1)),
         Mortality_Subadults2   = as.numeric(as.character(Mortality_Subadults2)),
         Mortality_Adults       = as.numeric(as.character(Mortality_Adults)),
         Repro                  = as.numeric(as.character(Repro)),
         N_Supplements          = as.numeric(as.character(N_Supplements)),
         Supplement_Time        = as.numeric(as.character(Supplement_Time)),
         Stoch_Frequency        = as.numeric(as.character(Stoch_Frequency)),
         Stoch_Severity         = as.numeric(as.character(Stoch_Severity)),
         Year                   = as.numeric(as.character(Year)),
         IndAll                 = as.numeric(as.character(IndAll)),
         IndB                   = as.numeric(as.character(IndB)),
         IndK                   = as.numeric(as.character(IndK)),
         IndUb                  = as.numeric(as.character(IndUb)),
         IndFP                  = as.numeric(as.character(IndFP)),
         IndP                   = as.numeric(as.character(IndP)),
         IndBFP                 = as.numeric(as.character(IndBFP)),
         IndKFP                 = as.numeric(as.character(IndKFP)),
         IndUbFP                = as.numeric(as.character(IndUbFP)),
         IndBP                  = as.numeric(as.character(IndBP)),
         IndKP                  = as.numeric(as.character(IndKP)),
         IndUbP                 = as.numeric(as.character(IndUbP)),
         IndJuv                 = as.numeric(as.character(IndJuv)),
         IndSub1                = as.numeric(as.character(IndSub1)),
         IndSub2                = as.numeric(as.character(IndSub2)),
         IndAdu                 = as.numeric(as.character(IndAdu)))

str(df_StevS)
```

## Reorder rows
```{r}
df_StevS  %<>%
  arrange(Run, Year)%>%
  tibble::as_tibble()
```

## Reorder columns
```{r}
df_StevS %<>% 
  select(Run, Scenario, Year, 
         IndAll,IndP,
         IndJuv,IndSub1,IndSub2,IndAdu,
         Mortality_Juveniles,Mortality_Subadults1,Mortality_Subadults2,Mortality_Adults,Repro,
         Supplemented,N_Supplements,Supplement_Time, Stoch_events, Stoch_Frequency, Stoch_Severity,
         N_Juveniles,N_Subadults1,N_Subadults2,N_Adults, Over_Scenario)
df_StevS <- droplevels(df_StevS)
```

# Save data
```{r}
#saveRDS(object = df_StevS, file = here("output", "data-proc", "model-proc", "df_StevS.rds"))
```


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

