---
title: "PVA of the NBI: Analysis of Stochastic Event Scenarios - Analysis" ## name of your project
author: "Sinah Drenske, Viktoriia Radchuk, Stephanie Kramer-Schadt"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 9, fig.height = 6, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Research question:** (Q1) How will the NBI population develop in the next 50 years and what will be the extinction risk, given that the colonies will be either managed or left to themselves? (Q2) Will the developments in the two already established colonies Burghausen & Kuchl be different? 
* **Study area:** Austria, Germany and Italy
* **Data:** Results of the second NetLogo Model


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("here")
library("tidyverse")     ## data manipulation
library("magrittr")
```


# Data
```{r data}
df_StevS <- readRDS(file = here("output", "data-proc", "NetLogo-proc", "df_StevS.rds"))
```

# Data overview
## Minimum
```{r}
df_min_StevS <- df_StevS %>% group_by(Run,Scenario) %>% summarize_if(is.numeric, min) # lowest value per run
```

```{r}
range(df_min_StevS$IndAll) # range of the lowest values: 0-104
```

```{r}
df_min_StevS <- droplevels(df_min_StevS)
```

```{r}
# How often went the population extinct per scenario (100 runs)? TRUE = went extinct, FALSE = went not extinct
df_min_StevS_tab <- table(df_min_StevS$Scenario, df_min_StevS$IndAll < 1) 
table(df_min_StevS$IndAll < 1)
```

```{r}
# Have a look at the table
# There is one row per scenario with the numbers of scenarios that went extinct or not
head(df_min_StevS_tab)
tail(df_min_StevS_tab)
```

## Extinctions
df_min_CataS_tab has 2 columns with overall 1604 values (802*2). The second column ("TRUE") gives the number of runs per scenario (at most 100) that went extinct.
```{r}
ext <- df_min_StevS_tab[,2]
```

```{r}
df_ext_StevS <- df_min_StevS %>% group_by(Scenario) %>% summarize_if(is.numeric, min) # calculate the lowest value of all runs per scenario
df_ext_StevS$Extinct <- ext
df_ext_StevS <- droplevels(df_ext_StevS)
```

```{r}
table(df_ext_StevS$Extinct) # Table the number of runs that went extinct. In 727 scenarios of 820 scenarios the population never went extinct
#727/820  = 88.66 %
```

```{r}
table(df_ext_StevS$Extinct < 6 ) 
# Extinction probability: How often went the population extinct 5 times or less?
# In 748 of 820 scenarios the population went not extinct more than 5 times of 100 runs (Extinction probability <= 5%)
# 748/820 = 91.22 %
```

```{r}
rm(df_min_StevS_tab,ext)
```

## Maximum
```{r}
df_max_StevS <- df_StevS %>% group_by(Run,Scenario) %>% summarize_if(is.numeric, max) # highest value per run
```

```{r}
range(df_max_StevS$IndAll) # 89 7386
```

```{r}
df_max_StevS <- droplevels(df_max_StevS)
df_max_StevS_sc <- df_max_StevS %>% group_by(Scenario) %>% summarize_if(is.numeric, max) # highest value per scenario
range(df_max_StevS_sc$IndAll) # 89 7386
```

## Save data
```{r}
# saveRDS(object = df_ext_StevS, file = here("output", "data-proc", "NetLogo-proc", "df_ext_StevS.rds"))
# saveRDS(object = df_max_StevS_sc, file = here("output", "data-proc", "NetLogo-proc", "df_max_StevS_sc.rds"))
```

# Calculation of Lambda
Calculate the difference between the years per run, put NA for the first value of each run (because there you can not calculate a difference)
```{r}
df_StevS <- df_StevS %>%  group_by(Run) %>% mutate(Diff = c(NA,diff(IndAll)))
```

Calculate the ratio (quotient) between the difference (step before) and the number of individuals in the year before
```{r}
df_StevS <- df_StevS %>%  group_by(Run) %>% mutate(Quot = Diff / lag(IndAll,1))
```

Calculate the stochastic Lambda by add 1 to the ratio (step before). Now we have a Lambda per year per run
```{r}
df_StevS <- df_StevS %>%  group_by(Run) %>% mutate(Lambda_sto = Quot + 1)
```

```{r}
range(df_StevS$Lambda_sto, na.rm = T) # 0 2 
df_StevS <- df_StevS[, -c(37,38)]
```

## Save data
```{r}
#saveRDS(object = df_StevS, file = here("output", "data-proc", "NetLogo-proc", "df_StevS.rds"))
```

# Mean per scenario
At first we calculated mean values per run. E.g. there is one value per run for Lambda or for the number of individuals during this run for all years of this run.
```{r}
## Mean for each run
df_run_StevS <- df_StevS %>%  group_by(Run, Scenario) %>% summarize_if(is.numeric, mean, na.rm = TRUE)
# 82000 rows, 1 for each Run. 
```

Calculate the SD of the stochastic Lambda per scenario
```{r}
df_scenario_StevS <- df_run_StevS %>% group_by(Scenario) %>% mutate(SD = sd(Lambda_sto, na.rm = T))
```

Calculate the SD
```{r}
df_sc_StevS_sd <- df_scenario_StevS %>% group_by(Scenario) %>% summarize(SD = mean(SD, na.rm = T))
```

Mean for each scenario (Lambda per Scenario)
```{r}
df_sc_StevS_mean <- df_StevS %>% group_by(Over_Scenario,Scenario) %>% summarize_if(is.numeric, mean, na.rm = TRUE) # 820 rows
```

Combine SD and Extinction per scenario with the mean per scenario
```{r}
df_sc_StevS_mean$SD_Lambda <- df_sc_StevS_sd$SD
df_sc_StevS_mean$Extinct <- df_ext_StevS$Extinct
rm(df_sc_StevS_sd)
df_sc_StevS_mean <- droplevels(df_sc_StevS_mean)
```

## Overview of the data
```{r}
range(df_sc_StevS_mean$Lambda_sto) # 0.91-1.40

table(df_sc_StevS_mean$Lambda_sto > 1) 
# TRUE for 714 scenarios - positive population growth in 714 scenarios 

table(df_sc_StevS_mean$Extinct <= 5)
# Extinction probability is 5% or less in 748 scenarios 

table(df_sc_StevS_mean$Lambda_sto > 1 & df_sc_StevS_mean$Extinct <= 5) 
# Positive population growth and an extinction probability of 5% or less in 714 scenarios of 820 scenarios (87.07%)
```

```{r}
table(df_sc_StevS_mean$Stoch_Frequency, df_sc_StevS_mean$Stoch_Severity, df_sc_StevS_mean$Over_Scenario)
```

## Save data
```{r}
#saveRDS(object = df_sc_StevS_mean, file = here("output", "data-proc", "NetLogo-proc", "df_sc_StevS_mean.rds"))
```

# Analysis I: Scenarios with positive population growth
```{r}
df_StevS_lambda_pos <- subset(df_sc_StevS_mean, df_sc_StevS_mean$Lambda_sto > 1 & df_sc_StevS_mean$Extinct <= 5)
df_StevS_lambda_pos <- droplevels(df_StevS_lambda_pos)
```

Table the distributions between the different mortalities and reproduction rates

a) Overall
```{r}
table(df_sc_StevS_mean$Mortality_Juveniles)
# 0.20  0.30  0.36 
#  160   160   500   
table(df_sc_StevS_mean$Mortality_Subadults1)
# 0.08  0.19  0.26   
#  160   160   500
table(df_sc_StevS_mean$Mortality_Subadults2)
# 0.14  0.24  0.31 
#  160   160   500  
table(df_sc_StevS_mean$Mortality_Adults)
# 0.02  0.14  0.22 
#  240   240   340 
table(df_sc_StevS_mean$Repro)
# 0.53  0.58  0.66  1.06  1.41  3.97 
#  400    80    80    80    80   100
```

b) scenarios with positive population growth
```{r}
table(df_StevS_lambda_pos$Mortality_Juveniles)
# 0.20  0.30  0.36 
#  160   160   394   
table(df_StevS_lambda_pos$Mortality_Subadults1)
# 0.08  0.19  0.26   
#  160   160   394   
table(df_StevS_lambda_pos$Mortality_Subadults2)
# 0.14  0.24  0.31 
#  160   160   394   
table(df_StevS_lambda_pos$Mortality_Adults)
# 0.02  0.14  0.22 
#  240   214   260  
table(df_StevS_lambda_pos$Repro)
# 0.53  0.58  0.66  1.06  1.41  3.97 
#  294    80    80    80    80   100
```

## Save Data
```{r}
#saveRDS(object = df_StevS_lambda_pos, file = here("output", "data-proc", "NetLogo-proc", "df_StevS_lambda_pos.rds"))
```

# Analysis II: Only Baseline values and Improvements
Now without Status quo, All chicks and All chicks without supplements
```{r}
df_sc_StevS_mean2 <- df_sc_StevS_mean[!(df_sc_StevS_mean$Over_Scenario %in% c("Statusquo","All_chicks","All_chicks_Cata")), ]
df_sc_StevS_mean2 <- droplevels(df_sc_StevS_mean2)

df_StevS_lambda_sub <- df_StevS_lambda_pos[ !(df_StevS_lambda_pos$Over_Scenario %in% c("Statusquo", "All_chicks", "All_chicks_Cata")), ]
```

a) Overall
```{r}
table(df_sc_StevS_mean2$Mortality_Juveniles)
# 0.20  0.30  0.36 
#  160   160   320 
table(df_sc_StevS_mean2$Mortality_Subadults1)
# 0.08 0.19 0.26 
#  160  160  320
table(df_sc_StevS_mean2$Mortality_Subadults2)
#0 .14 0.24 0.31 
#  160  160  320
table(df_sc_StevS_mean2$Mortality_Adults)
# 0.02 0.14 0.22 
#  240  240  160   
table(df_sc_StevS_mean2$Repro)
# 0.53 0.58 0.66 1.06 
#  400   80   80   80
```

b) Scenarios with positive population growth
```{r}
table(df_StevS_lambda_sub$Mortality_Juveniles)
# 0.20  0.30  0.36 
#  160   160   214    
table(df_StevS_lambda_sub$Mortality_Subadults1)
# 0.08 0.19 0.26 
#  160  160  214   
table(df_StevS_lambda_sub$Mortality_Subadults2)
#0 .14 0.24 0.31 
#  160  160  214    
table(df_StevS_lambda_sub$Mortality_Adults)
# 0.02 0.14 0.22 
#  240  214   80   
table(df_StevS_lambda_sub$Repro)
# 0.53 0.58 0.66 1.06 
#  294   80   80   80
```

# Comparison of values
## Over_scenario

```{r}
table(df_StevS_lambda_pos$Over_Scenario)
```

## Number and time of supplements
```{r}
table(df_sc_StevS_mean$N_Supplements, df_sc_StevS_mean$Supplement_Time)
```

```{r}
table(df_StevS_lambda_pos$N_Supplements,df_StevS_lambda_pos$Supplement_Time)
```

## catastrophe frequency and severity

```{r}
table(df_sc_StevS_mean$Stoch_Frequency, df_sc_StevS_mean$Stoch_Severity)
```

```{r}
table(df_StevS_lambda_pos$Stoch_Frequency, df_StevS_lambda_pos$Stoch_Severity)
```

```{r}
table(df_StevS_lambda_pos$Stoch_Frequency, df_StevS_lambda_pos$Stoch_Severity,df_StevS_lambda_pos$Mortality_Juveniles)
```

***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

