---
title: "PVA of the NBI: Analysis of Stochastic Event Scenarios - Plots" ## name of your project
author: "Sinah Drenske"                     ## your name(s)
date: "`r Sys.Date()`"                  ## current date
output:
  rmdformats::readthedown:
    code_folding: hide                  ## hide or show code by default?
    toc_depth: 3                        ## 3-level TOC
---

```{r setup, include=FALSE}
## You can ignore this chunk in most cases
## If you want to modify chunk options, you can do it here for all chunks or
## add the options in the repsective chunk header, e.g. `{r, message = FALSE}`
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE,
                      fig.width = 9, fig.height = 6, dpi = 500, 
                      retina = 1, fig.showtext = TRUE)
```

* **Hypothesis:** We hypothesize that at the present time the northern bald ibis population can survive without further management and release. We predict that the observed demographic rates will ensure population growth and do not differ between the breeding colonies. 
* **Study area:** Austria, Germany and Italy
* **Data:** Results of the second NetLogo model and tables created in script 07b

In this script we create barplots to show the distribution of survival and reproduction values between the scenarios per combination of stochastic event frequency and severity and we show the distribution between the number of supplements and the years of supplementation for scenarios with positive population growth.

# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("here")
library("tidyverse")
library("magrittr")
library("ggplot2")
library("viridisLite")
library("gridExtra")
```


# Data

```{r data}
df_StevS_lambda_pos <- readRDS(file = here("output", "data-proc", "model-proc", "df_StevS_lambda_pos.rds"))
df_sc_StevS_mean <- readRDS(file = here("output", "data-proc", "model-proc", "df_sc_StevS_mean.rds"))
```

# Barplot of all scenarios with positive population growth
## Version 1: Absolute numbers
Count the individuals
a) Overall
```{r}
table(df_sc_StevS_mean$Mortality_Juveniles)
# 0.20  0.30  0.36 
#  160   160   500 
table(df_sc_StevS_mean$Mortality_Subadults1)
# 0.08 0.19 0.26 
#  160  160  500
table(df_sc_StevS_mean$Mortality_Subadults2)
#0 .14 0.24 0.31 
#  160  160  500
table(df_sc_StevS_mean$Mortality_Adults)
# 0.02 0.14 0.22 
#  240  240  340   
table(df_sc_StevS_mean$Repro)
# 0.53 0.58 0.66 1.06  1.41 2.88
#  400   80   80   80    80  100
```

b) Scenarios with positive population growth
```{r}
table(df_StevS_lambda_pos$Mortality_Juveniles)
# 0.20  0.30  0.36 
#  160   160   394    
table(df_StevS_lambda_pos$Mortality_Subadults1)
# 0.08 0.19 0.26 
#  160  160  394   
table(df_StevS_lambda_pos$Mortality_Subadults2)
#0 .14 0.24 0.31 
#  160  160  394    
table(df_StevS_lambda_pos$Mortality_Adults)
# 0.02 0.14 0.22 
#  240  214  260   
table(df_StevS_lambda_pos$Repro)
# 0.53 0.58 0.66 1.06  1.41  3.97
#  294   80   80   80    80   100
```

Create a matrix
```{r}
StevS_mort_RR_mat_all <- matrix(data = c(
  500,160,160,NA, # Baseline, 10%, 25%
  500,160,160,NA, # Baseline, 10%, 25%
  500,160,160,NA, # Baseline, 10%, 25%
  340,240,240, NA, # Baseline, 10%, 25%
  400,80,80,80,80,100 # Baseline, 10%, 25%, Status quo, All chicks
), byrow = F)
StevS_mort_RR_mat_all

StevS_mort_RR_mat <- matrix( data = c(
  394,160,160,NA,
  394,160,160,NA,
  394,160,160,NA,
  260,214,240,NA,
  294,80,80,80,80,100 
), byrow = F)
StevS_mort_RR_mat
```

Plot
```{r}
par(mar = c(5,6,1,1))
barplot(StevS_mort_RR_mat_all, beside = T, ylim = c(0,700), space = c(0,-0.3),  las = 1,
        col = "grey", cex.axis = 1.5, cex.lab = 1.5,  border = NA)
barplot(StevS_mort_RR_mat, beside = T, ylim = c(0,700), space = c(0,-0.3),  las = 1,
        col = c(viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), cividis(2, begin = 0.3, end = 0.9)), cex.axis = 1.5, cex.lab = 1.5, ylab = "", border = NA, add = T)
axis(1, at = c(1.2, 4.9,8.6,12.3,16.5), labels = c("","","","",""), cex.axis = 1.5)
axis(1, at = c(1.2, 4.9,8.6,12.3,16.5), labels = c("s1","s2","s3","s4","RR"), cex.axis = 1.5, line = 0.2, lwd = 0)
mtext("Count", side = 2, line = 4, cex = 1.5)
box()
```

## Version 2: Percentage
Calculate the percent of the used s1-s4 and reproduction values in scenarios with positive population growth
```{r}
StevS_mort_RR_mat_perc <- (StevS_mort_RR_mat/StevS_mort_RR_mat_all) * 100
StevS_mort_RR_mat_perc
```

Plot
```{r}
par(mar = c(4,8,1,1))
barplot(StevS_mort_RR_mat_perc, beside = T, ylim = c(0,100), space = c(0,-0.3),  las = 1,
        col = c(viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), cividis(2, begin = 0.3, end = 0.9)), cex.axis = 2.5, cex.lab = 3, ylab = "", border = NA)
axis(1, at = c(1.2, 5.1,9.2,13.2,18.7), labels = c("","","","",""), cex.axis = 2.5)
axis(1, at = c(1.2, 5.1,9.2,13.2,18.7), labels = c("s1","s2","s3","s4","RR"), cex.axis = 2.5, line = 0.2, lwd = 0)
mtext("% of used values in scenarios", side = 2, line = 6, cex = 3)
mtext("with positive population growth", side = 2, line = 4, cex = 3)
box()
text(x = -2.8, y = -5, labels = "b)", xpd=NA, cex = 3)
#784*467
# In figure description: Please consider that overall, survival values increased by 25% were used more often
```

## Save plot
```{r}
# png(filename = here("plots", "04_PVA", "Stoch_eve_Supplements_bar.png"), width = 1200, height = 772)
# 
# par(mar = c(4,8,1,1))
# barplot(StevS_mort_RR_mat_perc, beside = T, ylim = c(0,100), space = c(0,-0.3),  las = 1,
#         col = c(viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), cividis(2, begin = 0.3, end = 0.9)), cex.axis = 2.5, cex.lab = 3, ylab = "", border = NA)
# axis(1, at = c(1.2, 5.1,9.2,13.2,18.7), labels = c("","","","",""), cex.axis = 2.5)
# axis(1, at = c(1.2, 5.1,9.2,13.2,18.7), labels = c("s1","s2","s3","s4","RR"), cex.axis = 2.5, line = 0.2, lwd = 0)
# mtext("% of used values in scenarios", side = 2, line = 6, cex = 3)
# mtext("with positive population growth", side = 2, line = 4, cex = 3)
# box()
# text(x = -2.8, y = -5, labels = "b)", xpd=NA, cex = 3)
# 
# dev.off()
```

A combinded figure of the barplots for the Baseline and Improvement scenarios and the stochastic event and supplement scenarios
```{r}
# Before this you have to run Script 06c until row 128
```

```{r}
# png(filename = here("plots", "04_PVA", "combined_bar.png"), width = 1200, height = 772)
# par(mar = c(0.4,4.6,1,0) + 0.1,
#     mfrow= c(2,1),
#     oma = c(3, 4, 0, 0) + 0.1)
#     #mai = c(1, 0.9, 0, 0.1))
# 
# barplot(mort_RR_mat_perc, beside = T, ylim = c(0,100), space = c(0,-0.3),  las = 1,
#         col = viridis(4, end = 0.95), cex.axis = 2.5, cex.lab = 3, border = NA,
#         ylab = " ")
# mtext("% of used values in scenarios", side = 2, line = 6.7, cex = 3, adj = 2)
# mtext("with positive population growth", side = 2, line = 4.7, cex = 3, adj = 2)
# box()
# text(x = -0.7, y = 90, labels = "a)", xpd=NA, cex = 3)
# #text(x = -2.8, y = -5, labels = "positive", xpd=NA, cex = 3)
# 
# barplot(StevS_mort_RR_mat_perc, beside = T, ylim = c(0,100), space = c(0,-0.3),  las = 1,
#         col = c(viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), viridis(4, end = 0.95), cividis(2, begin = 0.3, end = 0.9)), cex.axis = 2.5, cex.lab = 3, ylab = "", border = NA)
# axis(1, at = c(1.2, 5.1,9.2,13.2,18.7), labels = c("","","","",""), cex.axis = 2.5, outer = F)
# axis(1, at = c(1.2, 5.1,9.2,13.2,18.7), labels = c("s1","s2","s3","s4","RR"), cex.axis = 2.5, line = 0.2, lwd = 0, outer = F)
# #mtext("% of used values in scenarios", side = 2, line = 6, cex = 3)
# #mtext("with positive population growth", side = 2, line = 4, cex = 3)
# box()
# text(x = -0.7, y = 90, labels = "b)", xpd=NA, cex = 3)
# dev.off()
```

# Stochastic event frequency and severity
```{r}
df_StevS_summ <- df_StevS_lambda_pos %>% 
  group_by(Stoch_Frequency, Stoch_Severity) %>% 
  summarise(Freq = n())
```

```{r}
table(df_sc_StevS_mean$Stoch_Frequency, df_sc_StevS_mean$Stoch_Severity)
table(df_StevS_summ$Freq)
sum(df_StevS_summ$Freq)
```

```{r}
df_StevS_summ$Freq_perc <- (df_StevS_summ$Freq/41)*100 # value of the number of scenarios with positive population growth per combination of stochastic event severity and frequency / the overall number of scenarios per combination of stochastic event frequency and severity * 100 for percentage
  
range(df_StevS_summ$Freq_perc) # 80.4878 90.2439, the range is not so wide
```


## Version 1: Absolute numbers
```{r}
# Basic Plot
ggplot(data = df_StevS_summ, aes(x = Stoch_Severity, 
                                 y = Stoch_Frequency, 
                                 fill = Freq)) + 
  geom_tile() + 
  # Colour and borders
  scale_fill_viridis_c(option = "viridis") + 
  geom_hline(yintercept = c(0.025, 0.075, 0.125, 0.175, 0.225), 
             colour = "white", 
             size = 1) + 
  geom_vline(xintercept = c(0.025, 0.075, 0.125, 0.175, 0.225, 0.275), 
             colour = "white", 
             size = 1) + 
  # Axis breaks
  scale_x_continuous(#name = "Stochastic event severity",  
                     breaks = c(0.05, 0.10, 0.15, 0.20, 0.25), expand = c(0, 0)) + 
  scale_y_continuous(#name = "Stochastic event frequency", 
                     breaks = c(0.05, 0.10, 0.15, 0.20), expand = c(0, 0)) + 
  # Theme
  theme(rect = element_blank(), 
        axis.text = element_text(size = 14), 
        axis.ticks = element_line(size = 1), 
        axis.ticks.length = unit(.25, "cm"), 
        axis.title = element_text(size = 16), 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14)) + 
  labs(x = "Stochastic event severity", y = "Stochastic event frequency", fill = "Frequency")
```

## Version 2: Percentage
```{r}
# Basic Plot
ggplot(data = df_StevS_summ, aes(x = Stoch_Severity, 
                                 y = Stoch_Frequency, 
                                 fill = Freq_perc)) + 
  geom_tile() + 
  # Colour and borders
  scale_fill_viridis_c(option = "viridis") + 
  geom_hline(yintercept = c(0.025, 0.075, 0.125, 0.175, 0.225), 
             colour = "white", 
             size = 1) + 
  geom_vline(xintercept = c(0.025, 0.075, 0.125, 0.175, 0.225, 0.275), 
             colour = "white", 
             size = 1) + 
  # Axis breaks
  scale_x_continuous(#name = "Stochastic event severity",  
                     breaks = c(0.05, 0.10, 0.15, 0.20, 0.25), expand = c(0, 0)) + 
  scale_y_continuous(#name = "Stochastic event frequency", 
                     breaks = c(0.05, 0.10, 0.15, 0.20), expand = c(0, 0)) + 
  # Theme
  theme(rect = element_blank(), 
        axis.text = element_text(size = 14), 
        axis.ticks = element_line(size = 1), 
        axis.ticks.length = unit(.25, "cm"), 
        axis.title = element_text(size = 16), 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14)) + 
  labs(x = "Stochastic event severity", y = "Stochastic event frequency", fill = "Percent") + 
  geom_text(mapping = aes(y = 0.05, x = 0.05), label = "a)", vjust = 4.8, hjust = 5, size = 6) + 
  coord_cartesian(clip = "off")
```


## Save plot
The plot with percent is saved
```{r}
# ggsave(filename = here("plots", "04_PVA", "Stoch_event_percent.png"),
#        width = 13, height = 8.2, units = "cm")
```

# Number of supplements & years of supplementation
```{r}
table(df_sc_StevS_mean$Supplement_Time, df_sc_StevS_mean$N_Supplements)
```

```{r}
df_Supp_summ <- df_StevS_lambda_pos %>% 
  group_by(N_Supplements, Supplement_Time) %>% 
  summarise(Freq = n())
df_Supp_summ
```

Remove the combination with 0 supplements
```{r}
df_Supp_summ <- df_Supp_summ[-1,]
df_Supp_summ

table(df_sc_StevS_mean$N_Supplements, df_sc_StevS_mean$Supplement_Time)
```


```{r}
df_Supp_summ$Freq_perc <- (df_Supp_summ$Freq/200) *100

range(df_Supp_summ$Freq_perc) # 85.5 87.5, very small range
```


```{r}
df_Supp_summ$cat <- c(1:4)

cols <- c("N_Supplements", "Supplement_Time",  "cat")
df_Supp_summ[cols] <- lapply(df_Supp_summ[cols], as.factor)
```


```{r}
# Plot with percent of scenarios
ggplot(data = df_Supp_summ, aes(x = Supplement_Time, 
                                y = N_Supplements, 
                                fill = Freq_perc)) + 
  geom_tile() + 
  # Colour and borders
  scale_fill_viridis_c(option = "viridis") + 
  geom_hline(yintercept = c(0.5, 1.5, 2.5, 3.5), 
             colour = "white", 
             size = 1, na.rm = T) + 
  geom_vline(xintercept = c(0.5, 1.5, 2.5), 
             colour = "white", 
             size = 1) + 
  # Axis breaks
  scale_x_discrete(#name = "Time of supplementation",  
                     breaks = c(4,7), expand = c(0, 0)) +
  scale_y_discrete(#name = "Number of supplements", 
                   breaks= c(0,15,30), expand = c(0,0)) + 
   # Theme
  theme(rect = element_blank(), 
        axis.text = element_text(size = 14), 
        axis.ticks = element_line(size = 1), 
        axis.ticks.length = unit(.25, "cm"), 
        axis.title = element_text(size = 16), 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14), legend.box.spacing = unit(2.5, units = "cm")) + 
  labs(x = "Year of supplementing", y = "Number of supplements", fill = "Percent")+ 
  geom_text(mapping = aes(y = 1, x = 1), label = "b)", vjust = 6.5, hjust = 5.8, size = 6) + 
  coord_cartesian(clip = "off")
```


## Save plot
```{r}
# ggsave(filename = here("plots", "04_PVA", "Supplements_percent.png"),
#        width = 13, height = 8, units = "cm")
```

***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>

